---
title: "Astrocytes subpopulation scRNA-seq conserved markers comparison"
author: "Ibrahim Saliu"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

setwd("C:/Users/ibrof/Documents/ibrahim/Astrocyte_ConservedMarker_comparison")

options(future.globals.maxSize = 8000 * 1024^8)

install.packages("remotes")

```{r}

library(ComplexHeatmap)
library(dplyr) # for mutate
library(Matrix)
library(RColorBrewer) # colorRampPalette(), brewer.pal
library(ggplot2) # qplot(), position_nudge(), geom_text()
library(cowplot) # for plot_grid
library(gplots) # for heatmap2
library(VennDiagram)
library(multtest)
library(metap)
library(Seurat)
library("stringr")
library('purrr')
library(Hmisc) # for correlation and correlation plot
library(corrplot)
library(UpSetR)
library(matrixStats)
library(ComplexHeatmap)
library(dplyr)
library(stringr)
```

DimPlot(Yamada_Neuron_FinalObj, reduction = "umap", label = TRUE )

if (!require("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("")

remotes::install_version("SeuratObject", "4.1.4", repos = c("https://satijalab.r-universe.dev", getOption("repos")))
remotes::install_version("Seurat", "4.4.0", repos = c("https://satijalab.r-universe.dev", getOption("repos")))

```{r}

```

install.packages("C:/Users/ibrof/Downloads/SeuratObject_4.1.4.tar.gz", repos = NULL, type="source")


# merge Feleke's ast6 and 7 subpopulations
options(future.globals.maxSize = 8000 * 1024^8)
DimPlot(Ast_FinalObj, reduction = "umap", label = TRUE)

DimPlot(Ast_FinalObj, reduction = "umap", label = TRUE)

#find new conserved marker
DefaultAssay(Ast_FinalObj) <- "SCT"

Conserved_cluster_markers <- list() # list starting subscript is 1, can't use 0
for (i in 0:(nlevels(Ast_FinalObj@active.ident)-1)) {
  Conserved_cluster_markers[[i+1]] <- FindConservedMarkers(data_integrated, assay = "SCT", slot = "data", test.use = "wilcox", min.pct = 0.2, ident.1 = i, grouping.var = "Genotype", verbose = FALSE, only.pos = TRUE, recorrect_umi = FALSE)
  
   # apply cutoff
    Conserved_cluster_markers[[i+1]] <- Conserved_cluster_markers[[i+1]][ which(Conserved_cluster_markers[[i+1]]$max_pval < 0.05), ]
}

#output marker genes for each cluster
for (n in 1:length(Conserved_cluster_markers)) {
  file_number = n-1
  outputfile <- paste("Astrocyte_merge_cluster_ConservedMarkerGenes_cluster_", file_number,".csv", sep="")
  write.csv(Conserved_cluster_markers[n], outputfile)
}


#know all the name of the sample 
table(Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Ast_FinalObj$SampleNo <- Ast_FinalObj@meta.data$orig.ident

#.svg format
svg("UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Ast_FinalObj, reduction = "umap", pt.size = 0.8, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)  +  theme(legend.title = element_text(face = "bold", size = 8)) + theme(legend.text = element_text(size = 6)))
dev.off()

# cluster by sample
svg("bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
p1 <- DimPlot(Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
p1
dev.off()

# cluster by disease
svg("byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
p1 <- DimPlot(Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "Genotype", cols = c("green", "red")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
p1
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Ast_FinalObj), Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 9), rep("AD", 10))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)

meltData <- melt(test_data, id=c("Genotype"))
meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "AD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-2"))

#cell distribution and sample distribution
svg("PFC_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
  scale_color_manual(values=c("green", "red")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ Genotype, data = meltData)
summary(res.aov_PercentCell)
TukeyHSD(res.aov_PercentCell, conf.level = 0.99)

# Cell distribution in each disease condition
svg("PFC_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
  geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
  scale_fill_manual(values = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))  + labs(fill='Cluster')
dev.off()


#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("PFC_AD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*400,        # 5 x 300 pixels
    height = 6*650,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =1, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=27), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("PFC_AD_Astrocyte_dim15_res0.25_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()


# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("PFC_AD_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
DotPlot(Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("PFC_AD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()

png("PFC_AD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()

png("PFC_AD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()


#Vln plot for all datasets
dataset_list <- readRDS("~/ibrahim/All_Ast_dataset_obj.rds")

# Check the length of the dataset list
length(dataset_list)

# Get the names of the datasets
dataset_names <- names(dataset_list)

# Check and print the 'Genotype' counts for each dataset
for (name in dataset_names) {
  dataset <- dataset_list[[name]]
  cat("Dataset Name:", name, "\n")
  cat("Genotype Counts:\n")
  print(table(dataset@meta.data$Genotype))
  cat("\n")
}

# Define a function to process each dataset
process_dataset <- function(dataset) {
  DefaultAssay(dataset) <- "RNA"
  dataset <- NormalizeData(dataset, verbose = FALSE)
  return(dataset)
}

# Apply the function to each dataset in the list
dataset_list <- lapply(dataset_list, process_dataset)

gene <- c('GPC5', 'ITM2C', 'TNC')

# Loop through each dataset to generate violin plots
for (dataset_name in dataset_names) {
  current_dataset <- dataset_list[[dataset_name]]
  
  # Generate colors for the plot based on cell identities
  colors_list <- c(
    ifelse(grepl("Ast-0", levels(Idents(current_dataset))), "coral3", ""),
    ifelse(grepl("Ast-1", levels(Idents(current_dataset))), "deepskyblue3", ""),
    ifelse(grepl("Ast-Int", levels(Idents(current_dataset))), "dodgerblue4", ""),
    ifelse(grepl("Ast-2", levels(Idents(current_dataset))), "goldenrod2", ""),
    ifelse(grepl("Ast-3", levels(Idents(current_dataset))), "gray47", ""),
    ifelse(grepl("Ast-4", levels(Idents(current_dataset))), "mediumturquoise", ""),
    ifelse(grepl("Ast-5", levels(Idents(current_dataset))), "mediumorchid3", ""),
    ifelse(grepl("Ast-6", levels(Idents(current_dataset))), "gold4", ""),
    ifelse(grepl("Ast-7", levels(Idents(current_dataset))), "forestgreen", "")
  )
  filtered_colors <- colors_list[nzchar(colors_list)]
  
  # Generate and save violin plot
  filename <- paste(dataset_name, "NewConserMarker_VlnPlot.png", sep = "_")
  png(filename, width = 4*500, height = 4*300, res = 300, pointsize = 5)
  
  plot <- VlnPlot(current_dataset, assay = "RNA", features = gene, pt.size = 0, ncol = 3, cols = filtered_colors) &  theme(plot.title = element_text(size = rel(2.9)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=30), axis.text.y=element_text( size=25)) 
  print(plot)
  
  dev.off()
}


```{r}

```

# ########################### sequencing depth ###########################################
#nFeature
median_genes_df <- do.call(rbind, lapply(names(obj_list), function(dataset_name) {
  obj <- obj_list[[dataset_name]]
  
  # Replace 'orig.ident' with your sample ID column name if different
  df <- data.frame(
    sample = obj$orig.ident,
    nFeature_RNA = obj$nFeature_RNA
  )
  
  # Calculate median nFeature_RNA per sample
  median_per_sample <- df %>%
    group_by(sample) %>%
    summarise(median_nFeature_RNA = median(nFeature_RNA)) %>%
    mutate(dataset = dataset_name)
  
  median_per_sample
})) 

# Reorder columns if you like
median_genes_df <- median_genes_df[, c("dataset", "sample", "median_nFeature_RNA")]
median_genes_df$dataset <- factor(median_genes_df$dataset, levels = unique(median_genes_df$dataset))


#nCount
median_umis_df <- do.call(rbind, lapply(names(obj_list), function(dataset_name) {
  obj <- obj_list[[dataset_name]]
  
  # Replace 'orig.ident' with your sample ID column name if different
  df <- data.frame(
    sample = obj$orig.ident,
    nCount_RNA = obj$nCount_RNA
  )
  
  # Calculate median nCount_RNA per sample
  median_per_sample <- df %>%
    group_by(sample) %>%
    summarise(median_nCount_RNA = median(nCount_RNA)) %>%
    mutate(dataset = dataset_name)
  
  median_per_sample
})) 

# Reorder columns if you like
median_umis_df <- median_umis_df[, c("dataset", "sample", "median_nCount_RNA")]
median_umis_df$dataset <- factor(median_umis_df$dataset, levels = unique(median_umis_df$dataset))

#total nuclei
total_nuclei_df <- do.call(rbind, lapply(names(obj_list), function(dataset_name) {
  obj <- obj_list[[dataset_name]]
  
  df <- data.frame(
    sample = obj$orig.ident  # change if sample ID column differs
  )
  
  total_per_sample <- df %>%
    group_by(sample) %>%
    summarise(total_nuclei = n()) %>%
    mutate(dataset = dataset_name)
  
  total_per_sample
}))

# Reorder columns
total_nuclei_df <- total_nuclei_df[, c("dataset", "sample", "total_nuclei")]
total_nuclei_df$dataset <- factor (total_nuclei_df$dataset, levels = unique(total_nuclei_df$dataset))

#Gene per nuclei
png("Dataset_Gene_Metric_ScaterPlot.png", width = 7*350, height = 8*400, res = 300, pointsize = 5)
ggplot(median_genes_df, aes(x= dataset, y=median_nFeature_RNA, group = dataset, color = dataset, fill = dataset)) +
    scale_color_manual(values=c("black", "black", "black", "black", "black", "black", "black",  "black")) + geom_boxplot() +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  scale_fill_manual(values=c("ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite")) +
  labs(x = "Genotype", y = "Median Gene per Nuclei") +
  theme(legend.position="none", axis.text=element_text(size=23), axis.title.x =element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=32), axis.text.y=element_text(size=27)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()


#Umi per nuclei
png("Dataset_UMI_Metric_ScaterPlot.png", width = 7*350, height = 8*400, res = 300, pointsize = 5)
ggplot(median_umis_df, aes(x= dataset, y=median_nCount_RNA, group = dataset, color = dataset, fill = dataset)) +
    scale_color_manual(values=c("black", "black", "black", "black", "black", "black", "black",  "black")) + geom_boxplot() +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  scale_fill_manual(values=c("ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite")) +
  labs(x = "Genotype", y = "Median UMI per Nuclei") +
  theme(legend.position="none", axis.text=element_text(size=23), axis.title.x =element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=32), axis.text.y=element_text(size=27)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#total nuclei per sample
#Umi per nuclei
png("Dataset_total_nuclei_Metric_ScaterPlot.png", width = 7*350, height = 8*400, res = 300, pointsize = 5)
ggplot(total_nuclei_df, aes(x= dataset, y=total_nuclei, group = dataset, color = dataset, fill = dataset)) +
    scale_color_manual(values=c("black", "black", "black", "black", "black", "black", "black",  "black")) + geom_boxplot() +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  scale_fill_manual(values=c("ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite", "ghostwhite")) +
  labs(x = "Genotype", y = "Number of Nuclei") +
  theme(legend.position="none", axis.text=element_text(size=30), axis.title.x =element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=32), axis.text.y=element_text(size=27)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()



```{}

C:/Users/ibrof/Documents/ibrahim

```

###########################################################
# make a list of all the average expression of astrocyte subpopulation for all data
Ast_ConservedMarker_List <- list(Xu_ADPD_puAst_Ast0, Xu_ADPD_puAst_Ast1, Xu_ADPD_puAst_Ast2, Lau_AD_pfcAst_Ast0, Lau_AD_pfcAst_Ast1, Lau_AD_pfcAst_Ast2, Grubman_AD_ecAst_Ast0, Grubman_AD_ecAst_Ast1, Grubman_AD_ecAst_Ast2, Feleke_LBD_accAst_Ast0, Feleke_LBD_accAst_Ast1, Feleke_LBD_accAst_Ast2, Lee_HD_puAst_Ast0, Lee_HD_puAst_Ast2, Zhou_NHD_ocAst_Ast0, Zhou_NHD_ocAst_Ast1, Zhou_NHD_ocAst_Ast2, Lee_HD_cdAst_Ast0, Lee_HD_cdAst_Ast2, Gabitto_SEA_mtgAst_Ast0, Gabitto_SEA_mtgAst_Ast1, Gabitto_SEA_mtgAst_AstInt, Gabitto_SEA_mtgAst_Ast2)

# remove Ast3 subpopulation (Absinta_Ast3_AvgExp, Lau_Ast3_AvgExp, Lee_Ast3_AvgExp, Lee_HD_cdAst_Ast5, Gabitto_SEA_mtgAst_Ast6, Gabitto_SEA_mtgAst_Ast7)
# ##############################################################

# merge all the average expression dataframe in the list
All_Ast_ConservedMarker <- Reduce(function(x, y) merge(x, y, all=TRUE), Ast_ConservedMarker_List)
rownames(All_Ast_ConservedMarker) <- All_Ast_ConservedMarker$Gene
All_Ast_ConservedMarker$Gene <- NULL
# 1331 X 23

library(dplyr)
#remove genes that do are not expressed in atleast 3 dataset
DataToPlot_Ast <- All_Ast_ConservedMarker[rowSums(!is.na(All_Ast_ConservedMarker)) >= 3, ]
DataToPlot_Ast[is.na(DataToPlot_Ast)] <- 0
# convert all element greater than zero to 1
DataToPlot_Ast <- DataToPlot_Ast %>% mutate_if(is.numeric, ~1 * (. > 0))

range(DataToPlot_Ast)
# 0.00000 89.11574

library(circlize)
col_fun = colorRamp2(c(0, 1), c("white", "mediumblue"))
#col_fun(seq(0, 63))

#define genes to be labelled
genes.to.label <- c("GPC5", "NRXN1", "APOE", "CST3", "CLU", "VIM", "SERPINA3", "S100B", "HSPB8", "GJA1", "DPP10", "GFAP", "VCAN", "CD44", "GAPDH", "MT3", "TNC", "VIM", "SLC38A1")

position.to.lable <- which(rownames(DataToPlot_Ast) %in% genes.to.label)
genes.to.label <- rownames(DataToPlot_Ast)[c(which(rownames(DataToPlot_Ast) %in% genes.to.label))]

# define color annotation 
ColumnAnnotation <- HeatmapAnnotation(Region = str_sub(sapply(strsplit(colnames(DataToPlot_Ast), "_"), function(x){x[2]})), 
                                      Subpopulation = str_sub(sapply(strsplit(colnames(DataToPlot_Ast), "_"), function(x) {x[3]})),
                              col = list(Region = c(pu = "lightslateblue", "acc" = "antiquewhite2", "ec" = "blue4", "pfc" ="blueviolet", "oc" =  "yellow", "cd" = "cadetblue", "mtg" = "red4"), 
                              Subpopulation = c("Ast0" = "coral3", "Ast1" = "deepskyblue3", "Ast2" = "goldenrod2", "AstInt" = "dodgerblue4")), annotation_legend_param = list(grid_height = unit(0.7, "cm"), grid_width = unit(5, "mm"), labels_gp = gpar(fontsize = 14, face = "bold"), title_gp = gpar(fontsize = 14, fontface = "bold")))

RowAnnotation = rowAnnotation(genes = anno_mark(at = position.to.lable, labels = genes.to.label ,  labels_gp = gpar(fontsize = 17)))

png("Astrocytes_ConservedMarker_shared_heatmap_Expressed_in_AtLeast3.png", width = 10*500, height = 8*500, res = 300, pointsize = 8)
Heatmap(as.matrix(DataToPlot_Ast), name = "Trend", column_title = "Shared Conserved Markers in subpopulation and regions", cluster_columns  = TRUE, column_names_rot = 70, column_names_gp = gpar(fontsize = 17), col = col_fun, show_row_names = FALSE, column_km_repeats = 100, column_km = 3, row_km =3, right_annotation = RowAnnotation, top_annotation = ColumnAnnotation, rect_gp = gpar(col = "white", lwd = 0), heatmap_legend_param = list(labels_gp = gpar(fontsize = 12), legend_height = unit(5, "cm"), height = unit(2, "cm"), title_gp = gpar(fontsize = 14)))
dev.off()

svg("Astrocytes_ConservedMarker_shared_heatmap_Expressed_AtLeast3222.svg", width = 12,  height = 9,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
Heatmap(as.matrix(DataToPlot_Ast), name = "Trend", column_title = "Shared Conserved Markers in subpopulation and regions", cluster_columns  = TRUE, column_names_rot = 70, column_names_gp = gpar(fontsize = 17), col = col_fun, show_row_names = FALSE, column_km_repeats = 100, column_km = 3, row_km =3, right_annotation = RowAnnotation, top_annotation = ColumnAnnotation, rect_gp = gpar(col = "white", lwd = 0), heatmap_legend_param = list(labels_gp = gpar(fontsize = 12), legend_height = unit(5, "cm"), height = unit(2, "cm"), title_gp = gpar(fontsize = 14, fontface = "bold")))
dev.off()

pdf("Astrocytes_ConservedMarker_shared_heatmap_Expressed_AtLeast3.pdf", width = 14, height = 14)
Heatmap(as.matrix(DataToPlot_Ast), name = "Trend", column_title = "Shared Conserved Markers in subpopulation and regions", cluster_columns  = TRUE, column_names_rot = 70, column_names_gp = gpar(fontsize = 17), col = col_fun, show_row_names = FALSE, column_km_repeats = 100, column_km = 3, row_km =3, right_annotation = RowAnnotation, top_annotation = ColumnAnnotation, rect_gp = gpar(col = "white", lwd = 0), heatmap_legend_param = list(labels_gp = gpar(fontsize = 12), legend_height = unit(5, "cm"), height = unit(2, "cm"), title_gp = gpar(fontsize = 14)))
dev.off()


```{r}

```

# ################ Define common gene set ######################################

# ######## Subset each subpopulation and find the uniue genes in at least 5 dataset #####################
#### For Ast-0 ####
# List of marker sets
marker_sets <- c("Grubman_ecAst0", "Feleke_accAst0", "Lau_pfcAst0", "Xu_puAst0", "Zhou_ocAst0", "Lee_puAst0", "Lee_cdAst0", "Gabitto_SEA_mtgAst0")

# Create an empty list to store the data frames
result_list <- list()

# Loop over marker sets
for (marker_set in marker_sets) {
  # Create data frame
  current_df <- data.frame(Name = ConservedMarker_List[[marker_set]], Trend = 1)
  rownames(current_df) <- ConservedMarker_List[[marker_set]]
 
  # Append to the result list
  result_list[[marker_set]] <- current_df
}

colnames(result_list[["Grubman_ecAst0"]])[colnames(result_list[["Grubman_ecAst0"]]) == "Trend"] <- ""
colnames(result_list[["Feleke_accAst0"]])[colnames(result_list[["Feleke_accAst0"]]) == "Trend"] <- "RF_acc"
colnames(result_list[["Lau_pfcAst0"]])[colnames(result_list[["Lau_pfcAst0"]]) == "Trend"] <- "SL_pfc"
colnames(result_list[["Xu_puAst0"]])[colnames(result_list[["Xu_puAst0"]]) == "Trend"] <- "JX_pu"
colnames(result_list[["Zhou_ocAst0"]])[colnames(result_list[["Zhou_ocAst0"]]) == "Trend"] <- "YZ_oc"
colnames(result_list[["Lee_puAst0"]])[colnames(result_list[["Lee_puAst0"]]) == "Trend"] <- "HL_pu"
colnames(result_list[["Lee_cdAst0"]])[colnames(result_list[["Lee_cdAst0"]]) == "Trend"] <- "HL_cd"
colnames(result_list[["Gabitto_SEA_mtgAst0"]])[colnames(result_list[["Gabitto_SEA_mtgAst0"]]) == "Trend"] <- "MG_mtg"

# merge all the average expression dataframe in the list
df_ConservedMarker <- Reduce(function(x, y) merge(x, y, all=TRUE), result_list)
rownames(df_ConservedMarker) <- df_ConservedMarker$Name
df_ConservedMarker$Name <- NULL

# ###################### filter out all genes not expressed in 5 or more dataset
df_ConservedMarker <- df_ConservedMarker[rowSums(!is.na(df_ConservedMarker)) >= 5, ]
df_ConservedMarker[is.na(df_ConservedMarker)] <- "NotPresent"
df_ConservedMarker <- df_ConservedMarker %>% mutate_all(~ifelse(. == 1, "Present", .))
write.csv(df_ConservedMarker, file = "Astrocytes_subpopulation_genes_common_in_5_or_more_dataset_common_gene_set_Ast0.csv")

Ast0_common_genes <- rownames(df_ConservedMarker)

#### For Ast-1 ####
# List of marker sets
marker_sets <- c("Grubman_ecAst1", "Feleke_accAst1", "Lau_pfcAst1", "Xu_puAst1", "Zhou_ocAst1", "Gabitto_SEA_mtgAst1")

# Create an empty list to store the data frames
result_list <- list()

# Loop over marker sets
for (marker_set in marker_sets) {
  # Create data frame
  current_df <- data.frame(Name = ConservedMarker_List[[marker_set]], Trend = 1)
  rownames(current_df) <- ConservedMarker_List[[marker_set]]
 
  # Append to the result list
  result_list[[marker_set]] <- current_df
}

colnames(result_list[["Grubman_ecAst1"]])[colnames(result_list[["Grubman_ecAst1"]]) == "Trend"] <- "AG_ec"
colnames(result_list[["Feleke_accAst1"]])[colnames(result_list[["Feleke_accAst1"]]) == "Trend"] <- "RF_acc"
colnames(result_list[["Lau_pfcAst1"]])[colnames(result_list[["Lau_pfcAst1"]]) == "Trend"] <- "SL_pfc"
colnames(result_list[["Xu_puAst1"]])[colnames(result_list[["Xu_puAst1"]]) == "Trend"] <- "JX_pu"
colnames(result_list[["Zhou_ocAst1"]])[colnames(result_list[["Zhou_ocAst1"]]) == "Trend"] <- "YZ_oc"
colnames(result_list[["Gabitto_SEA_mtgAst1"]])[colnames(result_list[["Gabitto_SEA_mtgAst1"]]) == "Trend"] <- "MG_mtg"

# merge all the average expression dataframe in the list
df_ConservedMarker <- Reduce(function(x, y) merge(x, y, all=TRUE), result_list)
rownames(df_ConservedMarker) <- df_ConservedMarker$Name
df_ConservedMarker$Name <- NULL

# ###################### filter out all genes not expressed in 5 or more dataset
df_ConservedMarker <- df_ConservedMarker[rowSums(!is.na(df_ConservedMarker)) >= 5, ]
df_ConservedMarker[is.na(df_ConservedMarker)] <- "NotPresent"
df_ConservedMarker <- df_ConservedMarker %>% mutate_all(~ifelse(. == 1, "Present", .))

write.csv(df_ConservedMarker, file = "Astrocytes_subpopulation_genes_common_in_5_or_more_dataset_common_gene_set_Ast1.csv")
Ast1_common_genes <- rownames(df_ConservedMarker)

#### For Ast-2 ####
# List of marker sets
marker_sets <- c("Grubman_ecAst2", "Feleke_accAst2", "Lau_pfcAst2", "Xu_puAst2", "Zhou_ocAst2", "Lee_puAst2", "Lee_cdAst2", "Gabitto_SEA_mtgAst2")

# Create an empty list to store the data frames
result_list <- list()

# Loop over marker sets
for (marker_set in marker_sets) {
  # Create data frame
  current_df <- data.frame(Name = ConservedMarker_List[[marker_set]], Trend = 1)
  rownames(current_df) <- ConservedMarker_List[[marker_set]]
 
  # Append to the result list
  result_list[[marker_set]] <- current_df
}

colnames(result_list[["Grubman_ecAst2"]])[colnames(result_list[["Grubman_ecAst2"]]) == "Trend"] <- "AG_ec"
colnames(result_list[["Feleke_accAst2"]])[colnames(result_list[["Feleke_accAst2"]]) == "Trend"] <- "RF_acc"
colnames(result_list[["Lau_pfcAst2"]])[colnames(result_list[["Lau_pfcAst2"]]) == "Trend"] <- "SL_pfc"
colnames(result_list[["Xu_puAst2"]])[colnames(result_list[["Xu_puAst2"]]) == "Trend"] <- "JX_pu"
colnames(result_list[["Zhou_ocAst2"]])[colnames(result_list[["Zhou_ocAst2"]]) == "Trend"] <- "YZ_oc"
colnames(result_list[["Lee_puAst2"]])[colnames(result_list[["Lee_puAst2"]]) == "Trend"] <- "HL_pu"
colnames(result_list[["Lee_cdAst2"]])[colnames(result_list[["Lee_cdAst2"]]) == "Trend"] <- "HL_cd"
colnames(result_list[["Gabitto_SEA_mtgAst2"]])[colnames(result_list[["Gabitto_SEA_mtgAst2"]]) == "Trend"] <- "MG_mtg"

# merge all the average expression dataframe in the list
df_ConservedMarker <- Reduce(function(x, y) merge(x, y, all=TRUE), result_list)
rownames(df_ConservedMarker) <- df_ConservedMarker$Name
df_ConservedMarker$Name <- NULL

# ###################### filter out all genes not expressed in 5 or more dataset
df_ConservedMarker <- df_ConservedMarker[rowSums(!is.na(df_ConservedMarker)) >= 5, ]
df_ConservedMarker[is.na(df_ConservedMarker)] <- "NotPresent"
df_ConservedMarker <- df_ConservedMarker %>% mutate_all(~ifelse(. == 1, "Present", .))

write.csv(df_ConservedMarker, file = "Astrocytes_subpopulation_genes_common_in_5_or_more_dataset_common_gene_set_Ast2.csv")
Ast2_common_genes <- rownames(df_ConservedMarker)

# ##################### Matrix plot for consensus gene set Z-score #########################
#normalize data
#average expression
#z-score

dataset_obj <- readRDS("~/ibrahim/All_Ast_dataset_obj.rds")

#normalize data
dataset_obj <- lapply(dataset_obj, NormalizeData)

names(dataset_obj) <- c("AG_ec", "RF_acc", "SL_pfc", "JX_pu", "YZ_oc", "HL_pu", "HL_cd", "MG_mtg")

#avg expression of normalized expression
Average_exp_datasets <- lapply(dataset_obj, function(obj) AverageExpression(obj, assays="RNA", slot="data", group.by="ident")$RNA)

#subset only the homologous Ast-subpopulation for each data
clusters <- c("Ast-0", "Ast-1", "Ast-2")

Average_exp_subset <- lapply(Average_exp_datasets, function(mat) {
  mat[, intersect(clusters, colnames(mat)), drop = FALSE]
})

#rename the column in each dataset
Average_exp_subset <- lapply(names(Average_exp_subset), function(nm) {
    mat <- Average_exp_subset[[nm]]
    colnames(mat) <- paste0(nm, "_", colnames(mat))
    mat
})
names(Average_exp_subset) <- names(dataset_obj)

#calculate z-score for each data and convert into a matrix
common_genes <- Reduce(intersect, lapply(Average_exp_subset, rownames))

aligned <- lapply(Average_exp_subset, function(mat) {
  mat[common_genes, , drop = FALSE]
})

combined <- do.call(cbind, aligned)

#calculate z-score
zscore_mat <- t(scale(t(combined)))

#order in alphabetical order
zscore_mat <- zscore_mat[order(rownames(zscore_mat)), ]

zscore_df <- as.data.frame(zscore_mat)

zscore_df <- zscore_df[, c(
  grep("Ast-0$", colnames(zscore_df), value = TRUE),
  grep("Ast-1$", colnames(zscore_df), value = TRUE),
  grep("Ast-2$", colnames(zscore_df), value = TRUE)
)]

All_common_genes <- genes_vector
# remove AC073941.1 and ADAMTS9-AS2 cos no Zscore

genes_vector <- c("ADGRB3","ARHGAP24","CABLES1","CACNB2","CADM2","FLRT2","GPC5","GPM6A","GRM3","HPSE2","NRXN1","PLCB1","RORA","ZNF98","AGT","APOE","AQP4","ATP1B2","CKB","CLU","CPE","CST3",
  "FTH1","FTL","GJA1","GLUL","GPR37L1","ITM2B","ITM2C","MT1E","MT2A","MT3","PSAP","PTN",
  "SPARCL1","TUBB2B","VIM","ABCC4", "AHNAK","ANGPT1","ARHGEF4",
  "ATP9A","CD44","CERS6","CRYAB","DAAM2","DCLK1","DCLK2","DGKB","DLGAP1","DPP10","DPP6",
  "DPYSL3","GALNT15","GFAP","GLIS3","GPC6","ID3","ITPKB","KAZN","L3MBTL4","LINC00609",
  "LINC01088","LINC01094","MAN1C1","MAPT","MIR4300HG","NAV2","NFASC","PLEKHA5","PLXDC2",
  "PPP1R12B","RHPN1","RNF19A","SLC24A4","SLC38A1","SMAD9","SNED1","SORBS1","ST6GAL1",
  "ST6GALNAC3","TNC","UBC","VCAN","WDR49")

zscore_df_subset <- zscore_df[genes_vector, ]

library(circlize)
col_fun = colorRamp2(c(-3, 0, 3), c("red", "white", "mediumblue"))
#col_fun(seq(0, 63))

#define genes to be labelled
genes.to.label <- c("ADGRB3", "GPC5", "NRXN1", "RORA", "CABLES1", "CADM2", "GPM6A", "PLCB1", "RORA", "APOE", "ATP1B2", "CST3", "CLU", "GLUL", "GJA1", "MT2A", "ITM2C", "GJA1", "CPE", "SPARCL1", "VIM", "DPP10", "GFAP", "VCAN", "CD44", "DCLK1", "ABCC4", "ARHGEF4", "GLIS3", "TNC", "PLEKHA5", "MAPT", "L3MBTL4", "SLC38A1", "SORBS1", "CRYAB", "NAV2", "RNF19A", "ITPKB")

position.to.lable <- which(rownames(zscore_df_subset) %in% genes.to.label)
genes.to.label <- rownames(zscore_df_subset)[c(which(rownames(zscore_df_subset) %in% genes.to.label))]

# define color annotation 
ColumnAnnotation <- HeatmapAnnotation(region = str_sub(sapply(strsplit(colnames(zscore_df_subset), "_"), function(x){x[2]})), 
                                      subpopulation = str_sub(sapply(strsplit(colnames(zscore_df_subset), "_"), function(x) {x[3]})),
                      col = list(region = c(pu = "lightslateblue", "acc" = "antiquewhite2", "ec" = "blue4", "pfc" ="blueviolet", "oc" =  "yellow", "cd" = "cadetblue", "mtg" = "red4"),
                                subpopulation = c("Ast-0" = "coral3", "Ast-1" = "deepskyblue3", "Ast-2" = "goldenrod2")), annotation_name_gp = gpar(fontsize = 13, fontface = "bold"), annotation_legend_param = list(grid_height = unit(0.7, "cm"), grid_width = unit(6, "mm"), labels_gp = gpar(fontsize = 13), title_gp = gpar(fontsize = 15, fontface = "bold")))

RowAnnotation = rowAnnotation(genes = anno_mark(at = position.to.lable, labels = genes.to.label, labels_gp = gpar(fontsize = 11)))

png("2026_04_11_Astrocytes_consensus_ConservedMarker_geneset.png", width = 8*800, height = 10*1000, res = 300, pointsize = 8)  
Heatmap(as.matrix(zscore_df_subset), name = "Avg expression", cluster_rows = FALSE, cluster_columns = FALSE, column_names_rot = 84, column_names_gp = gpar(fontsize = 23), col = col_fun, show_row_names = FALSE, top_annotation = ColumnAnnotation, right_annotation = RowAnnotation, rect_gp = gpar(col = "white", lwd = 0), heatmap_legend_param = list(labels_gp = gpar(fontsize = 13), legend_height = unit(5, "cm"), legend_width = unit(7, "cm"), height = unit(5, "cm"), title_gp = gpar(fontsize = 15, fontface = "bold")))
dev.off()

save.image("Astrocytes_subpopulation_scRNA-seq_conserved_markerss_comparison.RData")



```{r}

```

# ################################## Conserved Marker Gene Ontology ##########################

# convert SYMBOL to ENTREZID in our data and Smjic data, universal parameter
# create a list of all the genes both in our data and Smjic 
All_Universe_Genes <- list(rownames(Xu_all_expressed_genes), rownames(Grubman_all_expressed_genes), rownames(Lau_all_expressed_genes), rownames(Feleke_all_expressed_genes), rownames(Lee_Pu_all_expressed_genes), rownames(Lee_Cd_all_expressed_genes), rownames(Zhou_all_expressed_genes), rownames(Gabitto_all_expressed_genes))

# convert
All_Universe_Genes_ENTREZID <- lapply(All_Universe_Genes, function(x) bitr(x, fromType = "SYMBOL",  toType = c("ENTREZID"),  OrgDb = org.Hs.eg.db))

All_Universe_Genes_ENTREZID <- lapply(All_Universe_Genes_ENTREZID, function(x) x$ENTREZID)

#make our conserved gene to list
All_ConservedMarker_genes <- list(JX_pu_Ast0 = ConsMarker_Xu_Pu_ADPD_Ast0$X, JX_pu_Ast1 = ConsMarker_Xu_Pu_ADPD_Ast1$Gene, JX_pu_Ast2 = ConsMarker_Xu_Pu_ADPD_Ast2$Gene, SL_pfc_Ast0 = ConsMarker_Lau_PFC_AD_Ast0$X, SL_pfc_Ast1 = ConsMarker_Lau_PFC_AD_Ast1$X, SL_pfc_Ast2 = ConsMarker_Lau_PFC_AD_Ast2$X, SL_pfc_Ast3 = ConsMarker_Lau_PFC_AD_Ast3$X, AG_ec_Ast0 = ConsMarker_Grubman_EC_AD_Ast0$X, AG_ec_Ast1 = ConsMarker_Grubman_EC_AD_Ast1$X, AG_ec_Ast2 = ConsMarker_Grubman_EC_AD_Ast2$X, RF_acc_Ast0 = ConsMarker_Feleke_ACC_LBD_Ast0$X, RF_acc_Ast1 = ConsMarker_Feleke_ACC_LBD_Ast1$X, RF_acc_Ast2 = ConsMarker_Feleke_ACC_LBD_Ast2$X, HL_cd_Ast0 = ConsMarker_Lee_Cd_HD_Ast0$X, HL_cd_Ast2 = ConsMarker_Lee_Cd_HD_Ast2$X, HL_cd_Ast5 = ConsMarker_Lee_Cd_HD_Ast5$X, HL_pu_Ast0 = ConsMarker_Lee_Pu_HD_Ast0$X, HL_pu_Ast2 = ConsMarker_Lee_Pu_HD_Ast2$X, HL_pu_Ast4 = ConsMarker_Lee_Pu_HD_Ast4$X, YZ_oc_Ast0 = ConsMarker_Zhou_oc_NHD_Ast0$X, YZ_oc_Ast1 = ConsMarker_Zhou_oc_NHD_Ast1$X, YZ_oc_Ast2 = ConsMarker_Zhou_oc_NHD_Ast2$X, MG_mtg_Ast0 = ConsMarker_Gabitto_mtg_AD_Ast0$X, MG_mtg_Ast1 = ConsMarker_Gabitto_mtg_AD_Ast1$X, MG_mtg_AstInt = ConsMarker_Gabitto_mtg_AD_AstInt$X, MG_mtg_Ast2 = ConsMarker_Gabitto_mtg_AD_Ast2$X, MG_mtg_Ast6 = ConsMarker_Gabitto_mtg_AD_Ast6$X, MG_mtg_Ast7 = ConsMarker_Gabitto_mtg_AD_Ast7$X)
                                  
#reorder the elements in the list
All_ConservedMarker_genes <- All_ConservedMarker_genes[c("AG_ec_Ast0", "RF_acc_Ast0", "SL_pfc_Ast0", "JX_pu_Ast0", "YZ_oc_Ast0", "HL_pu_Ast0", "HL_cd_Ast0",  "MG_mtg_Ast0", "AG_ec_Ast1", "RF_acc_Ast1", "SL_pfc_Ast1", "JX_pu_Ast1", "YZ_oc_Ast1", "MG_mtg_Ast1", "AG_ec_Ast2", "RF_acc_Ast2", "SL_pfc_Ast2", "JX_pu_Ast2", "YZ_oc_Ast2", "HL_pu_Ast2", "HL_cd_Ast2", "MG_mtg_Ast2", "MG_mtg_AstInt", "SL_pfc_Ast3", "HL_pu_Ast4", "HL_cd_Ast5", "MG_mtg_Ast6", "MG_mtg_Ast7")]

# convert SYMBOL to ENTREZID
ConservedMarker_List_ENTREZID <- lapply(All_ConservedMarker_genes, function(x) bitr(x, fromType = "SYMBOL",  toType = c("ENTREZID"),  OrgDb = org.Hs.eg.db))

ConservedMarker_List_ENTREZID <- lapply(ConservedMarker_List_ENTREZID, function(x) x$ENTREZID)


# Biological Process (BP)
Ast_compGO <- compareCluster(
  geneCluster   = ConservedMarker_List_ENTREZID,
  fun           = "enrichGO",
  ont = "BP", #CC #MF
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

Ast_ConsMarker_GO <- gsfilter(Ast_compGO, by = 'Count', min = 5)

write.csv( as.data.frame(Ast_ConsMarker_GO), "All_ConservedMarkerGene_compGO.csv")

Ast_ConsMarker_GO@compareClusterResult[["Gene_Ratio"]] <- sapply(strsplit(as.character(Ast_ConsMarker_GO@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

png("ConsMarker_GO3_New.png", width = 10*570, height = 8*500, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(Ast_ConsMarker_GO, showCategory = 4, title = "Biological Process", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(angle = 45,vjust=1, hjust=1), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold")) +    scale_x_discrete(drop=FALSE)
dev.off()

# ############################### KEGG #########################################
Ast_compKEGG <- compareCluster(geneCluster= ConservedMarker_List_ENTREZID,
                            fun           = "enrichKEGG",
                            organism = 'hsa', # mmu for mouse, hsa for human
                            universe = All_Universe_Genes_ENTREZID,
                            pvalueCutoff  = 0.05,
                            use_internal_data=F,
                            pAdjustMethod = "BH",
                            qvalueCutoff = 0.05)
 
Ast_ConsMarker_KEGG <- gsfilter(Ast_compKEGG, by = 'Count', min = 5)

Ast_ConsMarker_KEGG@compareClusterResult[["Gene_Ratio"]] <- sapply(strsplit(as.character(Ast_ConsMarker_KEGG@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

png("Ast_ConservedMarker_KEGG5.png", width = 12*500, height = 12*300, res = 300, pointsize = 4) 
plot <- clusterProfiler::dotplot(Ast_ConsMarker_KEGG, showCategory = 5, title = "KEGG pathway", label_format = 80, font.size = 22, by = "GeneRatio")
plot + theme(axis.text.x = element_text(angle = 45,vjust=1, hjust=1), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold")) +  scale_x_discrete(drop=FALSE)
dev.off()

# ############################## Venn Diagram for collagen-containing extracellular matrix ##########################
ConserMarker_CC <- read.csv("C:/Users/ibrof/Documents/ibrahim/Astrocyte_ConservedMarker_comparison/Conserved_Marker_GO_Analysis/All_ConservedMarkerGene_compGO_CC.csv")

df_ConserMarker_CC_CCEM <- ConserMarker_CC[ConserMarker_CC$Description =="collagen-containing extracellular matrix", ]


#collagen-containing extracellular matrix
df_ConserMarker_CC_CCEM_Ast1 <- df_ConserMarker_CC_CCEM[grepl("_Ast1", df_ConserMarker_CC_CCEM$Cluster), ]
df_ConserMarker_CC_CCEM_Ast1_marker <- unlist(strsplit(df_ConserMarker_CC_CCEM_Ast1$geneID, "/"))

df_ConserMarker_CC_CCEM_Ast2 <- df_ConserMarker_CC_CCEM[grepl("_Ast2", df_ConserMarker_CC_CCEM$Cluster), ]
df_ConserMarker_CC_CCEM_Ast2_marker <- unlist(strsplit(df_ConserMarker_CC_CCEM_Ast2$geneID, "/"))

# Make venn diagram
CCEM_genes <- list(Ast1 = df_ConserMarker_CC_CCEM_Ast1_marker, Ast2 = df_ConserMarker_CC_CCEM_Ast2_marker)

venn.diagram(CCEM_genes, filename="collagen-containing extracellular matrix_comparison2.png", imagetype = "png", height = 3*400, width = 3*400,  resolution = 300, fill=c("deepskyblue3", "goldenrod2"), scaled = FALSE, col = "transparent",
	alpha = 0.50,
	cex = 2.5,
	fontfamily = "serif",
	fontface = "bold",
	cat.col = c("black"),
	cat.cex = 2,
	cat.pos = c(0, 0),
	cat.dist = c(0.025, 0.025),
	cat.fontfamily = "serif",
	rotation.degree = 180,
	euler.d	= T)



