---
title: "2025_10_20_combine_UpDown_DEG_GO_AND DEG_ranking"
author: "Ibrahim Saliu"
date: "2026-01-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Seurat)
library(sctransform)
library(dplyr) # for mutate
library(RColorBrewer) # colorRampPalette(), brewer.pal
library(ggplot2) # qplot(), position_nudge(), geom_text()
library(cowplot) # for plot_grid
library(gplots) # for heatmap2
library(metap)
library("ggrepel") # for Volcano plot
library(clusterProfiler)
library(VennDiagram)
#library(org.Mm.eg.db)
library(org.Hs.eg.db)
library(Hmisc) # for capitalize
library(DOSE)
library(KEGG.db)
library(SCENIC)
library(UpSetR)
library(ComplexHeatmap)
library(stringr)
library(reshape)
library(ComplexUpset)

```

#read workspace and continue analysis
load("~/ibrahim/snRNA_seq_DEG_analysis/GO_DEG_analysis/Combine_all_Ast_subpopulation_for_GO_analysis_and_extract_terms/2025_08_30_All_data_Ast_subpopulation_combined_DEG_GO_analysis_workspace.RData")

#All upregulated DEGs list
DEG_Ast_upregulated <- list(YZ_ocNHD_Ast0 = ZY_ocNHD_Ast0_up, YZ_ocNHD_Ast1 = ZY_ocNHD_Ast1_up, YZ_ocNHD_Ast2 = ZY_ocNHD_Ast2_up, JX_puAD_Ast0 = XJ_puAD_Ast0_up, JX_puAD_Ast1 = XJ_puAD_Ast1_up, JX_puAD_Ast2 = XJ_puAD_Ast2_up, JX_puPD_Ast0 = XJ_puPD_Ast0_up, JX_puPD_Ast1 = XJ_puPD_Ast1_up, JX_puPD_Ast2 = XJ_puPD_Ast2_up, HL_cdHD_Ast0 = LH_cdHD_Ast0_up, HL_cdHD_Ast2 = LH_cdHD_Ast2_up, HL_cdHD_Ast5 = LH_cdHD_Ast5_up, HL_puHD_Ast0 = LH_puHD_Ast0_up, HL_puHD_Ast2 = LH_puHD_Ast2_up, HL_puHD_Ast4 = LH_puHD_Ast4_up, RF_accDLBD_Ast0 = FR_accDLBD_Ast0_up, RF_accDLBD_Ast1 = FR_accDLBD_Ast1_up, RF_accDLBD_Ast2 = FR_accDLBD_Ast2_up, RF_accPDD_Ast0 = FR_accPDD_Ast0_up, RF_accPDD_Ast1 = FR_accPDD_Ast1_up, RF_accPDD_Ast2 = FR_accPDD_Ast2_up, RF_accPD_Ast0 = FR_accPD_Ast0_up, RF_accPD_Ast1 = FR_accPD_Ast1_up, RF_accPD_Ast2 = FR_accPD_Ast2_up, SL_pfcAD_Ast0 = LS_pfcAD_Ast0_up, SL_pfcAD_Ast1 = LS_pfcAD_Ast1_up, SL_pfcAD_Ast2 = LS_pfcAD_Ast2_up, SL_pfcAD_Ast3 = LS_pfcAD_Ast3_up,  AG_ecAD_Ast0 = GA_ecAD_Ast0_up, AG_ecAD_Ast1 = GA_ecAD_Ast1_up, AG_ecAD_Ast2 = GA_ecAD_Ast2_up, MG_mtgDEM_Ast0 = GM_mtgAD_Ast0_up, MG_mtgDEM_Ast1 = GM_mtgAD_Ast1_up, MG_mtgDEM_Ast2 = GM_mtgAD_Ast2_up, MG_mtgDEM_Ast6 = GM_mtgAD_Ast6_up, MG_mtgDEM_Ast7 = GM_mtgAD_Ast7_up)

#All downregulated DEG list
DEG_Ast_downregulated <- list(YZ_ocNHD_Ast0 = ZY_ocNHD_Ast0_down, YZ_ocNHD_Ast1 = ZY_ocNHD_Ast1_down, YZ_ocNHD_Ast2 = ZY_ocNHD_Ast2_down, JX_puAD_Ast0 = XJ_puAD_Ast0_down, JX_puAD_Ast1 = XJ_puAD_Ast1_down, JX_puAD_Ast2 = XJ_puAD_Ast2_down, JX_puPD_Ast0 = XJ_puPD_Ast0_down, JX_puPD_Ast1 = XJ_puPD_Ast1_down, JX_puPD_Ast2 = XJ_puPD_Ast2_down, HL_cdHD_Ast0 = LH_cdHD_Ast0_down, HL_cdHD_Ast2 = LH_cdHD_Ast2_down, HL_cdHD_Ast5 = LH_cdHD_Ast5_down, HL_puHD_Ast0 = LH_puHD_Ast0_down, HL_puHD_Ast2 = LH_puHD_Ast2_down, HL_puHD_Ast4 = LH_puHD_Ast4_down, RF_accDLBD_Ast0 = FR_accDLBD_Ast0_down, RF_accDLBD_Ast1 = FR_accDLBD_Ast1_down, RF_accDLBD_Ast2 = FR_accDLBD_Ast2_down, RF_accPDD_Ast0 = FR_accPDD_Ast0_down, RF_accPDD_Ast1 = FR_accPDD_Ast1_down, RF_accPDD_Ast2 = FR_accPDD_Ast2_down, RF_accPD_Ast0 = FR_accPD_Ast0_down, RF_accPD_Ast1 = FR_accPD_Ast1_down, RF_accPD_Ast2 = FR_accPD_Ast2_down, SL_pfcAD_Ast0 = LS_pfcAD_Ast0_down, SL_pfcAD_Ast1 = LS_pfcAD_Ast1_down, SL_pfcAD_Ast2 = LS_pfcAD_Ast2_down, SL_pfcAD_Ast3 = LS_pfcAD_Ast3_down,  AG_ecAD_Ast0 = GA_ecAD_Ast0_down, AG_ecAD_Ast1 = GA_ecAD_Ast1_down, AG_ecAD_Ast2 = GA_ecAD_Ast2_down, MG_mtgDEM_Ast0 = GM_mtgAD_Ast0_down, MG_mtgDEM_Ast1 = GM_mtgAD_Ast1_down, MG_mtgDEM_Ast2 = GM_mtgAD_Ast2_down, MG_mtgDEM_Ast6 = GM_mtgAD_Ast6_down, MG_mtgDEM_Ast7 = GM_mtgAD_Ast7_down)


# ############ GO for upregulated and downregulated separately ################
# upregulated
DEG_Ast_up_ENTREZID <- lapply(DEG_Ast_upregulated, function(x) bitr(x, fromType = "SYMBOL",  toType = c("ENTREZID"),  OrgDb = org.Hs.eg.db))
DEG_Ast_up_ENTREZID <- lapply(DEG_Ast_up_ENTREZID, function(x) x$ENTREZID)

# downregulated
DEG_Ast_down_ENTREZID <- lapply(DEG_Ast_downregulated, function(x) bitr(x, fromType = "SYMBOL",  toType = c("ENTREZID"),  OrgDb = org.Hs.eg.db))
DEG_Ast_down_ENTREZID <- lapply(DEG_Ast_down_ENTREZID, function(x) x$ENTREZID)


# ############# BP GO categories for upregulated DEG #################
# compare GO enrichment for all subpopulation combining upregulated
DEG_Ast_upregulated_compGO_BP <- compareCluster(
  geneCluster   = DEG_Ast_up_ENTREZID,
  fun           = "enrichGO",
  ont = "BP",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

DEG_Ast_upregulated_compGO_BP <- gsfilter(DEG_Ast_upregulated_compGO_BP, by = 'Count', min = 5)

DEG_Ast_upregulated_compGO_BP@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(DEG_Ast_upregulated_compGO_BP@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

df_DEG_Ast_upregulated_compGO_BP  <- as.data.frame(DEG_Ast_upregulated_compGO_BP)

png("2026_01_30_DEG_Ast_subpopulation_upregulated_compGO_BP_Top2.png", width = 12*1000, height = 9*1000, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(DEG_Ast_upregulated_compGO_BP, showCategory = 2, title = "Top 2 upregulated BP", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 35, angle = 55,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=25), legend.title=element_text(size=25), legend.key.size = unit(1.0, 'cm')) +  scale_x_discrete(drop=FALSE)
dev.off()

# ############# BP GO categories for DOWNregulated DEG #################
# compare GO enrichment for all subpopulation combining upregulated
DEG_Ast_downregulated_compGO_BP <- compareCluster(
  geneCluster   = DEG_Ast_down_ENTREZID,
  fun           = "enrichGO",
  ont = "BP",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

DEG_Ast_downregulated_compGO_BP <- gsfilter(DEG_Ast_downregulated_compGO_BP, by = 'Count', min = 5)

DEG_Ast_downregulated_compGO_BP@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(DEG_Ast_downregulated_compGO_BP@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

df_DEG_Ast_downregulated_compGO_BP  <- as.data.frame(DEG_Ast_downregulated_compGO_BP)

png("2026_01_30_DEG_Ast_subpopulation_downregulated_compGO_BP_Top2.png", width = 12*1000, height = 9*1000, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(DEG_Ast_downregulated_compGO_BP, showCategory = 2, title = "Top 2 upregulated BP", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 35, angle = 55,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=25), legend.title=element_text(size=25), legend.key.size = unit(1.0, 'cm')) +  scale_x_discrete(drop=FALSE)
dev.off()

#combine DEGs of up and down regulated DEGs from each dataset into one
#upregulated
YZ_ocNHD_DEG_UpDown <- c(ZY_ocNHD_DEG_up, ZY_ocNHD_DEG_down)
HL_puHD_DEG_UpDown <- c(LH_puHD_DEG_up, LH_puHD_DEG_down)
JX_puAD_DEG_UpDown <- c(XJ_puAD_DEG_up, XJ_puAD_DEG_down)
JX_puPD_DEG_UpDown <- c(XJ_puPD_DEG_up, XJ_puPD_DEG_down)
MG_mtgDEM_DEG_UpDown <- c(GM_mtgAD_DEG_up, GM_mtgAD_DEG_down)
RF_accDLBD_DEG_UpDown <- c(FR_accDLBD_DEG_up, FR_accDLBD_DEG_down) 
RF_accPD_DEG_UpDown <- c(FR_accPD_DEG_up, FR_accPD_DEG_down)
RF_accPDD_DEG_UpDown <- c(FR_accPDD_DEG_up, FR_accPDD_DEG_down)
SL_pfcAD_DEG_UpDown <- c(LS_pfcAD_DEG_up, LS_pfcAD_DEG_down)
AG_ecAD_DEG_UpDown <- c(GA_ecAD_DEG_up, GA_ecAD_DEG_down)
HL_cdHD_DEG_UpDown <- c(LH_cdHD_DEG_up, LH_cdHD_DEG_down)

#create a list of the dataset DEGs
DEG_Ast_UpDown_combine <- list(YZ_ocNHD = YZ_ocNHD_DEG_UpDown, JX_puAD = JX_puAD_DEG_UpDown, JX_puPD = JX_puPD_DEG_UpDown, HL_cdHD = HL_cdHD_DEG_UpDown, HL_puHD = HL_puHD_DEG_UpDown, RF_accDLBD = RF_accDLBD_DEG_UpDown, RF_accPDD = RF_accPDD_DEG_UpDown, RF_accPD = RF_accPD_DEG_UpDown, SL_pfcAD = SL_pfcAD_DEG_UpDown, AG_ecAD = AG_ecAD_DEG_UpDown, MG_mtgDEM = MG_mtgDEM_DEG_UpDown)

saveRDS(DEG_Ast_UpDown_combine, file = "DEG_set_List_Ast_UpDown_combine.rds")

```{}

#remove duplicated DEG in each DEG vector in the list
List_DEG_Ast_UpDown_combine <- lapply(DEG_Ast_UpDown_combine, unique)

#create a dataframe for the DEG vector and assign 1 to the 
List_DEG_Ast_UpDown_combine <- lapply(names(List_DEG_Ast_UpDown_combine), function(name) {
  vec <- List_DEG_Ast_UpDown_combine[[name]]
  df <- data.frame(Gene = vec, Value = rep(1, length(vec)))  # Create dataframe with Value and Gene columns
  colnames(df)[2] <- name  # Rename 'Value' column to the name of the vector
  df
})

# Assign names to the dataframes in the list
names(List_DEG_Ast_UpDown_combine) <- names(List_DEG_Ast_UpDown_combine)

# reduce the list into a single dataframe
df_DEG_Ast_UpDown_combine <- Reduce(function(x, y) merge(x, y, all=TRUE), List_DEG_Ast_UpDown_combine)

# create a template
temp_df_DEG_Ast_UpDown_combine <- df_DEG_Ast_UpDown_combine

# define rownames and remove gene column
rownames(temp_df_DEG_Ast_UpDown_combine) <- temp_df_DEG_Ast_UpDown_combine$Gene
temp_df_DEG_Ast_UpDown_combine$Gene <- NULL

# Turn all the 1 value to character "Yes"
temp_df_DEG_Ast_UpDown_combine <- temp_df_DEG_Ast_UpDown_combine %>% mutate_all(funs(case_when( . == 1 ~ "Yes", TRUE ~ as.character(.))))

# counts the number of non-NA value and put the result in a new column "counts"
temp_df_DEG_Ast_UpDown_combine$counts <- rowSums(!is.na(temp_df_DEG_Ast_UpDown_combine))

# reorder the datafram on decreasing order of counts
temp_df_DEG_Ast_UpDown_combine <- temp_df_DEG_Ast_UpDown_combine[order(-temp_df_DEG_Ast_UpDown_combine$counts), ]

saveRDS(df_DEG_Ast_UpDown_combine, file = "combined_subpopulation_Ast_DEG_across_all_dataset.rds")
write.csv(temp_df_DEG_Ast_UpDown_combine, file = "Ranking_combined_subpopulation_Ast_DEG_across_all_dataset.csv")

temp_df_DEG_Ast_UpDown_combine$counts <- NULL

# overlap of DEG with GWAS AD risk genes in the DEGs across all dataset
#AD risk genes
AD_risk_gene <- read.csv("C:/Users/ibrof/Documents/ibrahim/Reference_data/AD_RiskGenes_Combined_McQuade_Karch.csv")

# PD reisk genes
PD_risk_gene <- read.csv("C:/Users/ibrof/Documents/ibrahim/Reference_data/PD_RiskGenes_Combined_Deng_Kumar_Nalls.csv")

# AD risk genes in the DEGs across all dataset
# Filter the df based on DEGs in the DEG dataframe
Ast_DEG_AD_risk_gene_overlap <-temp_df_DEG_Ast_UpDown_combine[rownames(temp_df_DEG_Ast_UpDown_combine) %in% AD_risk_gene$Gene, ]

# counts the number of non-NA value and put the result in a new column "counts"
Ast_DEG_AD_risk_gene_overlap$counts <- rowSums(!is.na(Ast_DEG_AD_risk_gene_overlap))

# reorder the datafram on decreasing order of counts
Ast_DEG_AD_risk_gene_overlap <- Ast_DEG_AD_risk_gene_overlap[order(-Ast_DEG_AD_risk_gene_overlap$counts), ]

# PD risk genes in the DEGs across all dataset
# Filter the df based on DEGs in the DEG dataframe
Ast_DEG_PD_risk_gene_overlap <-temp_df_DEG_Ast_UpDown_combine[rownames(temp_df_DEG_Ast_UpDown_combine) %in% PD_risk_gene$Gene, ]

# counts the number of non-NA value and put the result in a new column "counts"
Ast_DEG_PD_risk_gene_overlap$counts <- rowSums(!is.na(Ast_DEG_PD_risk_gene_overlap))

# reorder the datafram on decreasing order of counts
Ast_DEG_PD_risk_gene_overlap <- Ast_DEG_PD_risk_gene_overlap[order(-Ast_DEG_PD_risk_gene_overlap$counts), ]

write.csv(Ast_DEG_AD_risk_gene_overlap, file = "GWAS_AD_risk_gene_overla_with_combined_subpopulation_Ast_DEG_across_all_dataset.csv")
write.csv(Ast_DEG_PD_risk_gene_overlap, file = "GWAS_PD_risk_gene_overla_with_combined_subpopulation_Ast_DEG_across_all_dataset.csv")

saveRDS(df_DEG_Ast_UpDown_combine, file = "combined_subpopulation_Ast_DEG_across_all_dataset.rds")

```

# convert SYMBOL to ENTREZID
temp_DEG_Ast_UpDown_combine_ENTREZID <- lapply(DEG_Ast_UpDown_combine, function(x) bitr(x, fromType = "SYMBOL",  toType = c("ENTREZID"),  OrgDb = org.Hs.eg.db))

DEG_Ast_UpDown_combine_ENTREZID <- lapply(temp_DEG_Ast_UpDown_combine_ENTREZID, function(x) x$ENTREZID)

# compare GO enrichment 
combined_DEG_Ast_compGO <- compareCluster(
  geneCluster   = DEG_Ast_UpDown_combine_ENTREZID,
  fun           = "enrichGO",
  ont = "ALL",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

combined_DEG_Ast_compGO <- gsfilter(combined_DEG_Ast_compGO, by = 'Count', min = 5)

combined_DEG_Ast_compGO@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(combined_DEG_Ast_compGO@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

df_combined_DEG_Ast_compGO  <- as.data.frame(combined_DEG_Ast_compGO)

#plot
png("2025_10_20_Ast_combined_DEG_Ast_compGO_Top5.png", width = 13*800, height = 12*1000, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(combined_DEG_Ast_compGO, showCategory = 5, title = "Top 3 pathway", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 26, angle = 35,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold")) +  scale_x_discrete(drop=FALSE)
dev.off()

```{r}

```


load("~/ibrahim/snRNA_seq_DEG_analysis/GO_DEG_analysis/Combine_all_Ast_subpopulation_for_GO_analysis_and_extract_terms/Combine_ranking_and_GO/2026_01_28_Ast_combined_DEG_Ast_ranking_and_GO_analysis_workspace.RData")

```{}

#create a list of downregulated DEGs
DEG_Ast_downregulated_combine <- list(YZ_ocNHD_down = ZY_ocNHD_DEG_down, JX_puAD_down = XJ_puAD_DEG_down, JX_puPD_down = XJ_puPD_DEG_down, HL_cdHD_down = LH_cdHD_DEG_down, HL_puHD_down = LH_puHD_DEG_down, RF_accDLBD_down = FR_accDLBD_DEG_down, RF_accPDD_down = FR_accPDD_DEG_down, RF_accPD_down = FR_accPD_DEG_down, SL_pfcAD_down = LS_pfcAD_DEG_down, AG_ecAD_down = GA_ecAD_DEG_down, MG_mtgDEM_down = GM_mtgAD_DEG_down)

DEG_Ast_upregulated_combine <- list(YZ_ocNHD_up = ZY_ocNHD_DEG_up, JX_puAD_up = XJ_puAD_DEG_up, JX_puPD_up = XJ_puPD_DEG_up, HL_cdHD_up = LH_cdHD_DEG_up, HL_puHD_up = LH_puHD_DEG_up, RF_accDLBD_up = FR_accDLBD_DEG_up, RF_accPDD_up = FR_accPDD_DEG_up, RF_accPD_up = FR_accPD_DEG_up, SL_pfcAD_up = LS_pfcAD_DEG_up, AG_ecAD_up = GA_ecAD_DEG_up, MG_mtgDEM_up = GM_mtgAD_DEG_up)

DEG_Ast_UpDown <- c(DEG_Ast_upregulated_combine, DEG_Ast_downregulated_combine)

temp_DEG_Ast_UpDown_ENTREZID <- lapply(DEG_Ast_UpDown, function(x) bitr(x, fromType = "SYMBOL",  toType = c("ENTREZID"),  OrgDb = org.Hs.eg.db))

DEG_Ast_UpDown_ENTREZID <- lapply(temp_DEG_Ast_UpDown_ENTREZID, function(x) x$ENTREZID)

# compare GO enrichment 
DEG_Ast_UpDown_compGO <- compareCluster(
  geneCluster   = DEG_Ast_UpDown_ENTREZID,
  fun           = "enrichGO",
  ont = "ALL",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

DEG_Ast_UpDown_compGO <- gsfilter(DEG_Ast_UpDown_compGO, by = 'Count', min = 5)

DEG_Ast_UpDown_compGO@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(DEG_Ast_UpDown_compGO@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

#plot
png("2026_01_29_Ast_Subpopulation_combined_DEG_compGO_Top2.png", width = 13*800, height = 7*900, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO, showCategory = 2, title = "Top 2 Gene ontology", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 29, angle = 55,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE)
dev.off()

svg("2026_01_29_Ast_Subpopulation_combined_DEG_compGO_Top2.svg", width = 16,  height = 12) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO, showCategory = 2, title = "Top 2 Gene ontology", label_format = 80, font.size = 24, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(angle = 60,vjust=1, hjust=1, size = 19), axis.text.y = element_text(size = 18), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE) + NoLegend()
dev.off()

write.csv(DEG_Ast_UpDown_compGO, file = "2026_01_31_Ast_GO_enriched_pathways_in_DEG_sets_in_each_dataset.csv")

```


# ################################
# select certain terms to plot
# DEG_Ast_regulated_ENTREZID_combined_DEG contains the enriched pathway for all DEGs merged for all subpopulations from the same data.

PatternOfInterest <- c("Wnt signaling pathway", "cell-cell signaling by wnt", "regulation of Wnt signaling pathway", "negative regulation of Wnt signaling pathway", "canonical Wnt signaling pathway", "regulation of canonical Wnt signaling pathway", "positive regulation of Wnt signaling pathway", "small GTPase mediated signal transduction", "regulation of small GTPase mediated signal transduction", "JNK cascade", "actin filament organization", "microtubule binding", "lamellipodium", "cell leading edge", "glutamatergic synapse", "postsynaptic specialization", "neuron to neuron synapse", "beta-catenin binding", "establishment or maintenance of cell polarity", "focal adhesion", "cell-substrate junction", "adherens junction", "filopodium", "regulation of trans-synaptic signaling", "collagen-containing extracellular matrix", "synaptic membrane")

# ##########################
# dont' need to change below except ploting
List_TermOfInterest <- list()

for (pat in  PatternOfInterest ) {
  temp <- grep(pattern = pat, x = as.data.frame(DEG_Ast_UpDown_compGO)$Description , value = TRUE)
  List_TermOfInterest <- append(List_TermOfInterest, temp)
}

TermOfInterest <- do.call("rbind", List_TermOfInterest)
TermOfInterest <- unique(as.data.frame(TermOfInterest)$V1)

temp <- as.data.frame(DEG_Ast_UpDown_compGO)
# to get all terms contain the pattern
#selectedTerm = temp[which (temp$Description %in% TermOfInterest),]

# to get terms only in PatternOfInterest
selectedTerm = temp[which (temp$Description %in% PatternOfInterest),]

# define as compareClusterResult object
selectedTerm_compareClusterResult = new("compareClusterResult", compareClusterResult = selectedTerm)

# ##################################

#dotplot for selected term without empty columns
# plot with empty columns
svg("2026_01_29_Ast_Subpopulation_combined_regulated_DEG_compGO_SelectedTerm2.svg", width = 13,  height = 12,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
plot <- clusterProfiler::dotplot(selectedTerm_compareClusterResult, showCategory = 20, title = "Top 20 Selecetd GO Terms", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 17, angle = 70,vjust=1, hjust=1), axis.text.y = element_text(size = 18), axis.title.x =element_blank(), plot.title = element_text(size=18, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=15), legend.key.size = unit(0.6, 'cm')) +  scale_x_discrete(drop=FALSE) + NoLegend ()
dev.off()

```{r}

```

# ############################## reorder he terms (on the x-axis) and plot ################################
temp_selectedTerm_compareClusterResult <- as.data.frame(selectedTerm)

# Get the terms the top enriched terms (based on increasing p.adjust) per cluster
list_terms_ranked.by.p.adjust <- list()
for (i in names(DEG_Ast_UpDown)){
  # Extract the corresponding data frame from temp_selectedTerm_compareClusterResult
  temp_table <- temp_selectedTerm_compareClusterResult[which(temp_selectedTerm_compareClusterResult$Cluster == i), ]
  # Order the data frame by p.adjust in increasing order
  temp_table <- temp_table[order(temp_table$p.adjust), ]
  # Extract the top 2 enriched terms for the cluster and store them in the list
  list_terms_ranked.by.p.adjust[[i]] <- temp_table$Description[1:27]
}

list_Descriptions <- unlist(list_terms_ranked.by.p.adjust) %>% unique
list_Descriptions <- list_Descriptions[!is.na(list_Descriptions)] #remove NA values

# Construct the data table for plotting 
# Convert GeneRatio from a string to decimal via parse and eval functions
temp_selectedTerm_compareClusterResult$Gene_Ratio <- NA
for (n in 1:nrow(temp_selectedTerm_compareClusterResult)){ 
  temp_selectedTerm_compareClusterResult[n,"Gene_Ratio"] <- parse(text = temp_selectedTerm_compareClusterResult[n, "GeneRatio"]) %>% eval
}

# Create rows for clusters which does not show any enrichment in the Top2 from prior plot (e.g. LH_pu_Ast4_HD)
temp_df <- data.frame(Cluster = rep(c("RF_accPDD_up", "RF_accPD_up", "SL_pfcAD_up", "RF_accPDD_down"), times = length(list_Descriptions)),
                      ONTOLOGY = rep(NA, times = length(list_Descriptions)),
                      ID = rep(NA, times = length(list_Descriptions)),
                      Description = list_Descriptions, 
                      GeneRatio = rep("0", times = length(list_Descriptions)),
                      BgRatio = rep("0", times = length(list_Descriptions)),
                      RichFactor = rep("0", times = length(list_Descriptions)),
                      FoldEnrichment = rep("0", times = length(list_Descriptions)),
                      zScore = rep("0", times = length(list_Descriptions)),
                      pvalue = rep(0, times = length(list_Descriptions)),
                      p.adjust = rep(0, times = length(list_Descriptions)),
                      qvalue = rep(0, times = length(list_Descriptions)),
                      geneID = rep(NA, times = length(list_Descriptions)),
                      Count = rep(0, times = length(list_Descriptions)),
                      Gene_Ratio = rep(NA, times = length(list_Descriptions)))

temp_selectedTerm_compareClusterResult <- rbind(temp_selectedTerm_compareClusterResult, temp_df)

table_row.nums <- c()
for(i in list_Descriptions){
  table_row.nums <- c(table_row.nums, which(temp_selectedTerm_compareClusterResult$Description == i))
}

#Assign another name to the list of description
Descriptions <- list_Descriptions
#write out the desired order levels of the variable in the list
Descriptions <- c("Wnt signaling pathway", "cell-cell signaling by wnt", "regulation of Wnt signaling pathway", "negative regulation of Wnt signaling pathway", "canonical Wnt signaling pathway", "regulation of canonical Wnt signaling pathway", "positive regulation of Wnt signaling pathway", "beta-catenin binding", "small GTPase mediated signal transduction", "regulation of small GTPase mediated signal transduction", "JNK cascade", "actin filament organization", "microtubule binding", "lamellipodium", "filopodium", "cell leading edge", "establishment or maintenance of cell polarity", "adherens junction", "focal adhesion", "collagen-containing extracellular matrix", "cell-substrate junction", "regulation of trans-synaptic signaling", "postsynaptic specialization", "neuron to neuron synapse", "glutamatergic synapse",  "synaptic membrane") 

#set the order levels of factor variable
Descriptions <- factor(Descriptions, levels = Descriptions)

# Specify the order in the Dot Plot via factor()
temp_selectedTerm_compareClusterResult$Description <- factor(temp_selectedTerm_compareClusterResult$Description, levels = rev(Descriptions))

# Reformatted DotPlot using SCENIC::doHeatMap to include all clusters in axis (even those with no enriched terms)
svg("2026_01_29_Ast_Subpopulation_combined_DEG_compGO_selected_ordered.svg", width = 16,  height = 11.8,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
dotHeatmap(temp_selectedTerm_compareClusterResult[table_row.nums,c("Cluster", "Description", "Gene_Ratio", "p.adjust")], var.y = "Description", var.x = "Cluster", var.col = "p.adjust", 
           col.low = "red2" , col.mid = "indianred2", col.high = "blue", #cerulean
           var.size = "Gene_Ratio", min.size = 4, max.size = 9) &
  theme(axis.text.x = element_text(angle = 70, size = 22, color = "black"), 
        axis.text.y = element_text(size = 23, color = "black"), legend.title=element_text(size=15), legend.key.size = unit(0.6, 'cm')) + NoLegend()
dev.off()

png("2026_01_29_Ast_Subpopulation_combined_regulated_DEG_compGO_SelectedTerm_orderedddddd.png", width = 12*450, height = 7*550, res = 300, pointsize = 4)
dotHeatmap(temp_selectedTerm_compareClusterResult[table_row.nums,c("Cluster", "Description", "Gene_Ratio", "p.adjust")], var.y = "Description", var.x = "Cluster", var.col = "p.adjust", 
           col.low = "red2" , col.mid = "indianred2", col.high = "blue", #cerulean
           var.size = "Gene_Ratio", min.size = 4, max.size = 9) &
  theme(axis.text.x = element_text(angle = 70, size = 22, color = "black"), 
        axis.text.y = element_text(size = 23, color = "black"), legend.title=element_text(size=15), legend.key.size = unit(0.6, 'cm'))
dev.off()

write.csv(temp_selectedTerm_compareClusterResult, file = "2026_01_31_Wnt_signaling_selected_terms_and_its_target_pathways_for_manuscript.csv")

```{}

# #####################################################
# compare Biological Process enrichment for all subpopulation combining up- and down-regulated
DEG_Ast_UpDown_compGO_BP <- compareCluster(
  geneCluster   = DEG_Ast_UpDown_ENTREZID,
  fun           = "enrichGO",
  ont = "BP",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

DEG_Ast_UpDown_compGO_BP <- gsfilter(DEG_Ast_UpDown_compGO_BP, by = 'Count', min = 5)

DEG_Ast_UpDown_compGO_BP@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(DEG_Ast_UpDown_compGO_BP@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

df_DEG_Ast_UpDown_compGO_BP  <- as.data.frame(DEG_Ast_UpDown_compGO_BP)

png("2026_01_29_Ast_Subpopulation_combined_DEG_BP_Top2.png", width = 11*800, height = 7*900, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO_BP, showCategory = 2, title = "Top 2 Biological Process", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 29, angle = 35,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE)
dev.off()

svglite("2026_01_29_Ast_Subpopulation_combined_DEG_BP_Top2.svg", width = 16.5,  height = 10.4) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO_BP, showCategory = 2, title = "Top 2 Biological Process", label_format = 80, font.size = 24, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(angle = 60,vjust=1, hjust=1, size = 21), axis.text.y = element_text(size = 21), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE)
dev.off()

# compare Cellular Component enrichment for all subpopulation combining up- and down-regulated
DEG_Ast_UpDown_compGO_CC <- compareCluster(
  geneCluster   = DEG_Ast_UpDown_ENTREZID,
  fun           = "enrichGO",
  ont = "CC",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

DEG_Ast_UpDown_compGO_CC <- gsfilter(DEG_Ast_UpDown_compGO_CC, by = 'Count', min = 5)

DEG_Ast_UpDown_compGO_CC@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(DEG_Ast_UpDown_compGO_CC@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

df_DEG_Ast_UpDown_compGO_CC  <- as.data.frame(DEG_Ast_UpDown_compGO_CC)

png("2026_01_29_Ast_Subpopulation_combined_DEG_CC_Top2.png", width = 11*800, height = 6*700, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO_CC, showCategory = 2, title = "Top 2 Cellular component", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 29, angle = 35,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE)
dev.off()

svglite("2026_01_29_Ast_Subpopulation_combined_DEG_CC_Top2.svg", width = 14.5,  height = 8) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO_CC, showCategory = 2, title = "Top 2 Cellular component", label_format = 80, font.size = 24, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(angle = 60,vjust=1, hjust=1, size = 19), axis.text.y = element_text(size = 20), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE)
dev.off()

# compare Molecular function enrichment for all subpopulation combining up- and down-regulated
DEG_Ast_UpDown_compGO_MF <- compareCluster(
  geneCluster   = DEG_Ast_UpDown_ENTREZID,
  fun           = "enrichGO",
  ont = "MF",
  universe = All_Universe_Genes_ENTREZID,
  pvalueCutoff  = 0.05,
  pAdjustMethod = "BH",
  qvalueCutoff = 0.05,
  OrgDb='org.Hs.eg.db', 
  readable = TRUE)

DEG_Ast_UpDown_compGO_MF <- gsfilter(DEG_Ast_UpDown_compGO_MF, by = 'Count', min = 5)

DEG_Ast_UpDown_compGO_MF@compareClusterResult$"Gene_Ratio" <- sapply(strsplit(as.character(DEG_Ast_UpDown_compGO_MF@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

df_DEG_Ast_UpDown_compGO_MF  <- as.data.frame(DEG_Ast_UpDown_compGO_MF)

png("2026_01_29_Ast_Subpopulation_combined_DEG_MF_Top2.png", width = 11*1000, height = 6*700, res = 300, pointsize = 9) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO_MF, showCategory = 2, title = "Top 2 Molecular function", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(size = 29, angle = 35,vjust=1, hjust=1), axis.text.y = element_text(size = 38), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE)
dev.off()

svg("2026_01_29_Ast_Subpopulation_combined_DEG_MF_Top2.svg", width = 16.5,  height = 8) 
plot <- clusterProfiler::dotplot(DEG_Ast_UpDown_compGO_MF, showCategory = 2, title = "Top 2 Molecular function", label_format = 80, font.size = 24, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(angle = 45,vjust=1, hjust=1, size = 19), axis.text.y = element_text(size = 18), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold"), legend.text=element_text(size=15), legend.title=element_text(size=18)) +    scale_x_discrete(drop=FALSE) + NoLegend()
dev.off()

# ####################################
Ast_DEG_KEGG <- compareCluster(geneCluster= DEG_Ast_UpDown_ENTREZID,
                            fun           = "enrichKEGG",
                            organism = 'hsa', # mmu for mouse, hsa for human
                            universe = All_Universe_Genes_ENTREZID,
                            pvalueCutoff  = 0.05,
                            use_internal_data=T,
                            pAdjustMethod = "BH",
                            qvalueCutoff = 0.05)

Ast_DEG_KEGG <- gsfilter(Ast_DEG_KEGG, by = 'Count', min = 5)

Ast_DEG_KEGG@compareClusterResult[["Gene_Ratio"]] <- sapply(strsplit(as.character(Ast_DEG_KEGG@compareClusterResult[["GeneRatio"]]), "/"), function(x) as.numeric(x[1]) / as.numeric(x[2]))

png("Ast_DEG_KEGG.png", width = 10*500, height = 8*600, res = 300, pointsize = 4) 
plot <- clusterProfiler::dotplot(Ast_DEG_KEGG, showCategory = 3, title = "KEGG pathway", label_format = 80, font.size = 22, by = "Gene_Ratio")
plot + theme(axis.text.x = element_text(angle = 60,vjust=1, hjust=1), axis.title.x =element_blank(), plot.title = element_text(size=30, face = "bold")) +  scale_x_discrete(drop=FALSE)
dev.off()

```

save.image("2026_01_29_Ast_combined_DEG_Ast_ranking_and_GO_analysis_workspace.RData")

```{r}

```


