---
title: "Plots from all astrocytes objects"
author: "Ibrahim Saliu"
date: "`r Sys.Date()`"
output: html_document
---

setwd("C:/ibrahim/Plots")

```{r setup, include=FALSE}

library('Seurat')
library(sctransform)
library(cowplot)
library(gplots) # for heatmap
library(RColorBrewer) # colorRampPalette(), brewer.pal
library(ggplot2) # qplot(), position_nudge(), geom_text()
library(dplyr)
library("limma")
library(VennDiagram)
library(reshape)
library("ggpubr")
library(UpSetR)
library(gplots) # for heatmap2
library(VennDiagram)
library(multtest)
library(metap)
library("ggrepel") # for Volcano plot
library(Hmisc) # for correlation and correlation plot
library(corrplot)
library(ComplexHeatmap)
library('purrr')

```

options(future.globals.maxSize = 8000 * 1024^8)

```{r}

#Lau data
Lau_Ast_FinalObj <- readRDS("C:/ibrahim/Astrocyte_Psuedotime_analysis/2020_Lau_Astrocyte_ClusterRenamed_dim15_res0.25_ObjFinal.rds")

#plot
p1 <- DimPlot(Lau_PFC_Ast_FinalObj, reduction = "umap", pt.size = 1, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(legend.text = element_text(size = 10)) + theme(legend.key.size = unit(0.4, 'cm'))
png("Lau_PFC_AD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()

#know all the name of the sample 
table(Lau_PFC_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Lau_PFC_Ast_FinalObj$SampleNo <- Lau_PFC_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Lau_PFC_Ast_FinalObj@meta.data$SampleNo <- ifelse(Lau_PFC_Ast_FinalObj@meta.data$SampleNo == "AD1", "AD_1", Lau_PFC_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "AD1" to "AD_1"

#.svg format
svg("Lau_UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lau_PFC_Ast_FinalObj, reduction = "umap", pt.size = 0.8, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)  +  theme(legend.title = element_text(face = "bold", size = 8)) + theme(legend.text = element_text(size = 6)))
dev.off()

# cluster by sample
svg("Lau_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
p1 <- DimPlot(Lau_PFC_Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
p1
dev.off()

# cluster by disease
svg("Lau_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
p1 <- DimPlot(Lau_PFC_Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "Genotype", cols = c("green", "red")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
p1
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Lau_PFC_Ast_FinalObj), Lau_PFC_Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 9), rep("AD", 10))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)

meltData <- melt(test_data, id=c("Genotype"))
meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "AD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-2", "Ast-3"))

#cell distribution and sample distribution
svg("Lau_PFC_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c("green", "red")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ Genotype, data = meltData)
summary(res.aov_PercentCell)
TukeyHSD(res.aov_PercentCell, conf.level = 0.99)

# Cell distribution in each disease condition
svg("Lau_PFC_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))  + labs(fill='Cluster')
dev.off()


#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Lau_PFC_AD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*400,        # 5 x 300 pixels
    height = 6*650,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Lau_PFC_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =1, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=27), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Lau_PFC_AD_Astrocyte_dim15_res0.25_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lau_PFC_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()


# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Lau_PFC_AD_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
  DotPlot(Lau_PFC_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Lau_PFC_AD_Astrocyte_dim15_res0.25_DotPlot.svg", width = 7,  height =2.3,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Lau_PFC_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()


#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Lau_PFC_AD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lau_PFC_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lau_PFC_AD_Astrocyte_dim15_res0.25_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lau_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Lau_PFC_AD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lau_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lau_PFC_AD_Astrocyte_dim15_res0.25_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lau_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Lau_PFC_AD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lau_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lau_PFC_AD_Astrocyte_dim15_res0.25_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lau_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()

```

#Grubman data
Grubman_Ast_FinalObj <- readRDS("C:/ibrahim/Astrocyte_Psuedotime_analysis/2019_Grubman_Astrocyte_ClusterRenamed_dim15_res0.25_ObjFinal.rds")

#plot
p1 <- DimPlot(Grubman_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Grubman_EC_AD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()

#.svg format
svg("Grubman_EC_AD_Astrocyte_dim15_res0.25_RenamedCluster.svg", width = 10,  height = 9,  onefile = TRUE, pointsize = 13, family = "Arial", antialias = "gray")
p1 <- DimPlot(Grubman_Ast_FinalObj, reduction = "umap", pt.size = 2, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
p1
dev.off()


#know all the name of the sample 
table(Grubman_EC_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Grubman_EC_Ast_FinalObj$SampleNo <- Grubman_EC_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Grubman_EC_Ast_FinalObj@meta.data$SampleNo <- ifelse(Grubman_EC_Ast_FinalObj@meta.data$SampleNo == "Ct5", "CTRL_3", Grubman_EC_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "Ct5" to "CTRL_3"

#.svg format
svg("Grubman_UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Grubman_EC_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "deepskyblue3", "goldenrod2"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)) & NoLegend()
dev.off()

# cluster by sample
svg("Grubman_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Grubman_EC_Ast_FinalObj, reduction = "umap", pt.size = 1.2, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# cluster by disease
svg("Grubman_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Grubman_EC_Ast_FinalObj, reduction = "umap", pt.size = 1.2, label = F, group.by = "Genotype", cols = c("green", "red")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Grubman_Ast_FinalObj), Grubman_Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 3), rep("AD", 3))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)
meltData <- melt(test_data, id=c("Genotype"))

# reformat the column in the melta dataframe
meltData$Genotype <- ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "AD", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-1" & meltData$Genotype == "AD", paste0(meltData$Genotype, "1"),
                      ifelse(meltData$variable == "Ast-1" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "1"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "AD", paste0(meltData$Genotype, "0"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "0"),
                      meltData$Genotype))))))

#meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "AD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-2"))

#cell distribution and sample distribution
svg("Grubman_EC_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c("green", "red")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ variable, data = meltData)
summary(res.aov_PercentCell)
TukeyHSD(res.aov_PercentCell, conf.level = 0.99)

# Cell distribution in each disease condition
svg("Grubman_EC_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "deepskyblue3", "goldenrod2"))  + labs(fill='Cluster')
dev.off()


#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Grubman_EC_AD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*400,        # 5 x 300 pixels
    height = 6*650,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Grubman_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =1, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=27), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Grubman_EC_AD_Astrocyte_dim15_res0.25_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Grubman_EC_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()


# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Grubman_EC_AD_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
  DotPlot(Grubman_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Grubman_EC_AD_Astrocyte_dim15_res0.25_DotPlot.svg", width = 7,  height =2,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Grubman_EC_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()


#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Grubman_EC_AD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Grubman_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Grubman_EC_AD_Astrocyte_dim15_res0.25_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Grubman_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Grubman_EC_AD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Grubman_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Grubman_EC_AD_Astrocyte_dim15_res0.25_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Grubman_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Grubman_EC_AD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Grubman_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Grubman_EC_AD_Astrocyte_dim15_res0.25_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Grubman_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()


```{r}

Feleke_Ast_FinalObj <- readRDS("C:/ibrahim/Astrocyte_Psuedotime_analysis/Human_2021_Feleke_ACC_Astrocyte_Pseudotime/Feleke_LBD_Astrocyte_MergedClusterRenamed_dim15_res0.25_CleanObj_Final.rds")

#renamed cluster and merge cluster 2 and 3 to Ast-2 instead of Ast 2.1 and 2.2
new.cluster.ids <- c( "Ast-0", "Ast-2",  "Ast-2", "Ast-1")
names(new.cluster.ids) <- levels(Feleke_Ast_FinalObj)
Feleke_Ast_FinalObj <- RenameIdents(Feleke_Ast_FinalObj, new.cluster.ids)

# set the order for CTRL, AD and PD
Feleke_Ast_FinalObj@meta.data$Genotype <- factor(Feleke_Ast_FinalObj@meta.data$Genotype,levels = c("CTRL", "DLBD", "PD", "PDD"))

# set the order for "Ast-0", "Ast-1",  "Ast-2"
Feleke_Ast_FinalObj@active.ident <- factor(Feleke_Ast_FinalObj@active.ident,levels = c("Ast-0", "Ast-1", "Ast-2"))

#Feleke data
p1 <- DimPlot(Feleke_Ast_FinalObj, reduction = "umap", pt.size = 1, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Feleke_ACC_AD_Astrocyte_dim15_res0.20_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()


saveRDS(Feleke_Ast_FinalObj, file= "C:/ibrahim/Astrocyte_Psuedotime_analysis/Human_2021_Feleke_ACC_Astrocyte_Pseudotime/Feleke_LBD_Astrocyte_MergedClusterRenamed_dim15_res0.25_CleanObj_Final.rds")

#know all the name of the sample 
table(Feleke_ACC_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Feleke_ACC_Ast_FinalObj$SampleNo <- Feleke_ACC_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Feleke_ACC_Ast_FinalObj@meta.data$SampleNo <- ifelse(Feleke_ACC_Ast_FinalObj@meta.data$SampleNo == "PD_4_ACC_PD666", "PD_4", Feleke_ACC_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "PD_4_ACC_PD666" to "PD_4"

#.svg format
svg("Feleke_UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Feleke_ACC_Ast_FinalObj, reduction = "umap", pt.size = 0.7, cols =c("coral3", "deepskyblue3", "goldenrod2"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)) & NoLegend()
dev.off()

# cluster by sample
svg("Feleke_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Feleke_ACC_Ast_FinalObj, reduction = "umap", pt.size = 0.6, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 12)) + theme(legend.key.size = unit(0.35, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# cluster by disease
svg("Feleke_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Feleke_ACC_Ast_FinalObj, reduction = "umap", pt.size = 0.6, label = F, group.by = "Genotype", cols = c("green",  "orange1", "blue", "purple2")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Feleke_ACC_Ast_FinalObj), Feleke_ACC_Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 7), rep("DLBD", 6), rep("PD", 7), rep("PDD", 7))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)

meltData <- melt(test_data, id=c("Genotype"))
meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "DLBD", "PD", "PDD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-2"))

#cell distribution and sample distribution
svg("Feleke_ACC_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 9,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c("green",  "orange1", "blue", "purple2")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=16), axis.text.x=element_text(size=16), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ Genotype, data = meltData)
summary(res.aov_PercentCell)
TukeyHSD(res.aov_PercentCell, conf.level = 0.99)

# Cell distribution in each disease condition
svg("Feleke_ACC_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "deepskyblue3", "goldenrod2"))  + labs(fill='Cluster')
dev.off()


#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Feleke_ACC_AD_Astrocyte_dim15_res0.20_Vln.png",      
    width = 4*400,        # 5 x 300 pixels
    height = 6*650,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Feleke_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =1, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=27), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Feleke_ACC_AD_Astrocyte_dim15_res0.20_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Feleke_ACC_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()

# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Feleke_ACC_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
  DotPlot(Feleke_ACC_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Feleke_ACC_Astrocyte_dim15_res0.25_DotPlot.svg", width = 7,  height =2,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Feleke_ACC_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()

#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Feleke_ACC_AD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Feleke_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Feleke_ACC_AD_Astrocyte_dim15_res0.25_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Feleke_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Feleke_ACC_AD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Feleke_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Feleke_ACC_AD_Astrocyte_dim15_res0.25_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Feleke_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Feleke_ACC_AD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Feleke_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Feleke_ACC_AD_Astrocyte_dim15_res0.25_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Feleke_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()

```

#Xu data
Xu_Ast_FinalObj <- readRDS("C:/ibrahim/Astrocyte_Psuedotime_analysis/Putamen_data_pseudotime_analysis/Putamen_Astrocyte_FinalObj_dim15_res0.25_ClusterRenamed.rds")

#plot
p1 <- DimPlot(Xu_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "deepskyblue3", "goldenrod2", "gray47"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Xu_PU_ADPD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()


#know all the name of the sample 
table(Xu_Pu_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Xu_Pu_Ast_FinalObj$SampleNo <- Xu_Pu_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Xu_Pu_Ast_FinalObj@meta.data$SampleNo <- ifelse(Xu_Pu_Ast_FinalObj@meta.data$SampleNo == "AD_Pu_12265", "AD_1", Xu_Pu_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "AD_Pu_12265" to "AD_1"

#.svg format
svg("Xu_UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Xu_Pu_Ast_FinalObj, reduction = "umap", pt.size = 1, cols =c("coral3", "deepskyblue3", "goldenrod2"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)) & NoLegend()
dev.off()

# cluster by sample
svg("Xu_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Xu_Pu_Ast_FinalObj, reduction = "umap", pt.size = 0.8, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# cluster by disease
svg("Xu_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Xu_Pu_Ast_FinalObj, reduction = "umap", pt.size = 0.8, label = F, group.by = "Genotype", cols = c("green", "red", "blue")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Xu_Ast_FinalObj), Xu_Ast_FinalObj$orig.ident), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 4), rep("AD", 4), rep("PD", 4))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)
meltData <- melt(test_data, id=c("Genotype"))

# reformat the column in the melta dataframe
meltData$Genotype <- ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "HD", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-1" & meltData$Genotype == "HD", paste0(meltData$Genotype, "1"),
                      ifelse(meltData$variable == "Ast-1" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "1"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "HD", paste0(meltData$Genotype, "0"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "0"),
                      meltData$Genotype)))))) 
                      
meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "AD", "PD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-2"))

#cell distribution and sample distribution
svg("Xu_Pu_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c("green", "red", "blue")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ variable, data = meltData)
summary(res.aov_PercentCell)
TukeyHSD(res.aov_PercentCell, conf.level = 0.99)

# Cell distribution in each disease condition
svg("Xu_Pu_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "deepskyblue3", "goldenrod2"))  + labs(fill='Cluster')
dev.off()

#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Xu_PU_ADPD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*400,        # 5 x 300 pixels
    height = 6*650,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Xu_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =1, cols = c("coral3", "deepskyblue3", "goldenrod2", "gray47"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=27), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Xu_PU_ADPD_Astrocyte_dim15_res0.25_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Xu_Pu_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()

# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Xu_PU_ADPD_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
  DotPlot(Xu_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Xu_PU_ADPD_Astrocyte_dim15_res0.25_DotPlot.svg", width = 7,  height =2,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Xu_Pu_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()

#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Xu_PU_ADPD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Xu_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei")  + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()

# .svg format
svg("Xu_PU_ADPD_Astrocyte_dim15_res0.25_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Xu_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Xu_PU_ADPD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Xu_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Xu_PU_ADPD_Astrocyte_dim15_res0.25_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Xu_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Xu_PU_ADPD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Xu_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Xu_PU_ADPD_Astrocyte_dim15_res0.25_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Xu_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()


```{r}

#Lee data
Lee_Pu_Ast_FinalObj <- readRDS("C:/ibrahim/Astrocyte_Psuedotime_analysis/Lee_Pu_HD_Astrocyte_dim15_res0.25_FinalObj.rds")

#plot
p1 <- DimPlot(Lee_Pu_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "goldenrod2", "mediumturquoise"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Lee_PUHD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off() 

#renamed cluster Ast-0, Ast-2, Ast-4
new.cluster.ids <- c( "Ast-0", "Ast-2",  "Ast-4")
names(new.cluster.ids) <- levels(Lee_Pu_Ast_FinalObj)
Lee_Pu_Ast_FinalObj <- RenameIdents(Lee_Pu_Ast_FinalObj, new.cluster.ids)

# set the order for CTRL, HD
Lee_Pu_Ast_FinalObj@meta.data$Genotype <- factor(Lee_Pu_Ast_FinalObj@meta.data$Genotype,levels = c("CTRL", "HD"))

# set the order for "Ast-0", "Ast-2",  "Ast-4"
Lee_Pu_Ast_FinalObj@active.ident <- factor(Lee_Pu_Ast_FinalObj@active.ident,levels = c("Ast-0", "Ast-2", "Ast-4"))

DimPlot(Lee_Pu_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "goldenrod2", "mediumturquoise"), label = T) 

#know all the name of the sample 
table(Lee_Pu_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Lee_Pu_Ast_FinalObj$SampleNo <- Lee_Pu_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Lee_Pu_Ast_FinalObj@meta.data$SampleNo <- ifelse(Lee_Pu_Ast_FinalObj@meta.data$SampleNo == "PD_4_ACC_PD666", "PD_4", Lee_Pu_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "Control_4294"  to  "CTRL_1"

#.svg format
svg("Lee_Pu_UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Pu_Ast_FinalObj, reduction = "umap", pt.size = 1.7, cols =c("coral3", "goldenrod2", "mediumturquoise"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)) & NoLegend()
dev.off()

svg("Lee_Pu_UMAP2.svg",  width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Pu_HD_Astrocyte_dim15_res0.25_FinalObj, reduction = "umap", pt.size = 1.0, cols =c("coral3", "goldenrod2", "mediumturquoise"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10) +  theme(legend.title = element_text(face = "bold", size = 8)) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm'))) 
dev.off()

# cluster by sample
svg("Lee_Pu_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Pu_Ast_FinalObj, reduction = "umap", pt.size = 1.2, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# cluster by disease
svg("Lee_Pu_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Pu_Ast_FinalObj, reduction = "umap", pt.size = 1.2, label = F, group.by = "Genotype", cols = c("green", "midnightblue")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Lee_Pu_HD_Ast_FinalObj), Lee_Pu_HD_Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 5), rep("HD", 6))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)
meltData <- melt(test_data, id=c("Genotype"))

# reformat the column in the melta dataframe
meltData$Genotype <- ifelse(meltData$variable == "Ast-4" & meltData$Genotype == "HD", paste0(meltData$Genotype, "4"),
                      ifelse(meltData$variable == "Ast-4" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "4"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "HD", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "HD", paste0(meltData$Genotype, "0"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "0"),
                      meltData$Genotype))))))

#meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "HD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-2", "Ast-4"))

#cell distribution and sample distribution
svg("Lee_Pu_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c( "green", "midnightblue")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ variable, data = meltData)
summary(res.aov_PercentCell)
#            Df Sum Sq Mean Sq F value Pr(>F)    
#Genotype     5  23801    4760   118.9 <2e-16 ***
#Residuals   27   1081      40                   
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

TukeyHSD(res.aov_PercentCell, conf.level = 0.99)
# all combination is significant except 
#                    diff         lwr        upr     p adj
# HD0 vs CTRL0    -4.5235404 -18.88205   9.834965 0.8420562
# HD2 vs CTRL2     4.3277963 -10.03071  18.686302 0.8647728
# HD4 vs CTRL4     0.1957441 -14.16276  14.554250 0.9999999

# Cell distribution in each disease condition
svg("Lee_Pu_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "goldenrod2", "mediumturquoise"))  + labs(fill='Cluster')
dev.off()


#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Lee_PUHD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*400,        # 5 x 300 pixels
    height = 6*650,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Lee_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "goldenrod2", "mediumturquoise"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=27), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Lee_PUHD_Astrocyte_dim15_res0.25_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Pu_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "goldenrod2", "mediumturquoise"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()

# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Lee_PUHD_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
  DotPlot(Lee_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Lee_PUHD_Astrocyte_dim15_res0.25_DotPlot.svg", width = 7,  height =2,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Lee_Pu_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()


#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Lee_PUHD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lee_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lee_PUHD_Astrocyte_dim15_res0.25_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Lee_PUHD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lee_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lee_PUHD_Astrocyte_dim15_res0.25_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei")  + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Lee_PUHD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lee_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lee_PUHD_Astrocyte_dim15_res0.25_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()

```

#Zhou's data
Zhou_Ast_FinalObj <- readRDS("C:/ibrahim/Human_2023_Zhou_OC_NHD_Colona/Human_2023_Zhou_OccipitalCortex_NHD_Astrocyte/Zhou_OC_NHD_Astrocyte_integrated_dim15_res0.25_RenamedCluster.rds")

#plot
p1 <- DimPlot(Zhou_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "deepskyblue3", "goldenrod2"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Zhou_NHD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
    width = 5*400,        # 5 x 300 pixels
    height = 4*400,
    res = 300,            # 300 pixels per inch
    pointsize = 8)        # smaller font size
p1
dev.off()

#know all the name of the sample 
table(Zhou_OC_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Zhou_OC_Ast_FinalObj$SampleNo <- Zhou_OC_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Zhou_OC_Ast_FinalObj@meta.data$SampleNo <- ifelse(Zhou_OC_Ast_FinalObj@meta.data$SampleNo == "Control1_merged", "CTRL_1M", Zhou_OC_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "Control1_merged" to "CTRL_1M"

#.svg format
svg("Zhou_UMAP.svg", width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Zhou_OC_Ast_FinalObj, reduction = "umap", pt.size = 1.2, cols =c("coral3", "deepskyblue3", "goldenrod2"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)) & NoLegend()
dev.off()

# cluster by sample
svg("Zhou_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Zhou_OC_Ast_FinalObj, reduction = "umap", pt.size = 0.8, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# cluster by disease
svg("Zhou_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Zhou_OC_Ast_FinalObj, reduction = "umap", pt.size = 0.8, label = F, group.by = "Genotype", cols = c("green", "sienna4")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()


# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Zhou_NHD_Ast_FinalObj), Zhou_NHD_Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 11), rep("NHD", 3))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)
meltData <- melt(test_data, id=c("Genotype"))

# reformat the column in the melta dataframe
meltData$Genotype <- ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "NHD", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-1" & meltData$Genotype == "NHD", paste0(meltData$Genotype, "1"),
                      ifelse(meltData$variable == "Ast-1" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "1"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "NHD", paste0(meltData$Genotype, "0"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "0"),
                      meltData$Genotype))))))

#meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "NHD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-2"))

#cell distribution and sample distribution
svg("Zhou_OC_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c("green", "sienna4")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()
  
#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ variable, data = meltData)
summary(res.aov_PercentCell)
#            Df Sum Sq Mean Sq F value   Pr(>F)    
#Genotype     5   1584   316.9    7.54 6.27e-05 ***
#Residuals   36   1513    42.0                     
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

TukeyHSD(res.aov_PercentCell, conf.level = 0.99)
#                   diff         lwr        upr     p adj
# CTRL1 vs CTRL0 -14.5789239 -24.6577850 -4.5000627 0.0000896
# NHD1 vs CTRL0  -15.8128821 -31.2085968 -0.4171673 0.0076627
# CTRL2 vs CTRL1   9.3807875  -0.6980737 19.4596487 0.0193832
# NHD2 vs CTRL1   12.9494670  -2.4462477 28.3451818 0.0435575

# Cell distribution in each disease condition
svg("Zhou_OC_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "deepskyblue3", "goldenrod2"))  + labs(fill='Cluster')
dev.off()


#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Zhou_NHD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*700,        # 5 x 300 pixels
    height = 4*350,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Zhou_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2"))& 
  theme(plot.title = element_text(size = rel(3)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=31), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Zhou_NHD_Astrocyte_dim15_res0.25_Vln.svg", width = 7, height = 4, onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Zhou_OC_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "goldenrod2"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()


# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Zhou_NHD_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
DotPlot(Zhou_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Zhou_NHD_Astrocyte_dim15_res0.25_DotPlot.svg", width = 7,  height =2,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Zhou_OC_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()

#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Zhou_OC_NHD_Astrocyte_dim15_res0.25_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Zhou_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Zhou_OC_NHD_Astrocyte_dim15_res0.25_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Zhou_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Zhou_OC_NHD_Astrocyte_dim15_res0.25_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Zhou_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Zhou_OC_NHD_Astrocyte_dim15_res0.25_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Zhou_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") +  theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Zhou_OC_NHD_Astrocyte_dim15_res0.25_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Zhou_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Zhou_OC_NHD_Astrocyte_dim15_res0.25_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Zhou_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "deepskyblue3", "goldenrod2")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()


```{r}

#Lee data
Lee_Cd_Ast_FinalObj <- readRDS("C:/ibrahim/Human_2020_Lee_Cd_HD_Astrocyte_data/Lee_Cd_HD_Astrocyte_dim15_res0.20_FinalObj_RenamedCluster.rds")

#plot
p1 <- DimPlot(Lee_Cd_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "goldenrod2", "gray47"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Lee_cdHD_Astrocyte_dim15_res0.20_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()

#renamed cluster Ast-0, Ast-2, Ast-5
new.cluster.ids <- c( "Ast-0", "Ast-2", "Ast-5")
names(new.cluster.ids) <- levels(Lee_Cd_Ast_FinalObj)
Lee_Cd_Ast_FinalObj <- RenameIdents(Lee_Cd_Ast_FinalObj, new.cluster.ids)

# set the order for CTRL, HD
Lee_Cd_Ast_FinalObj@meta.data$Genotype <- factor(Lee_Cd_Ast_FinalObj@meta.data$Genotype,levels = c("CTRL", "HD"))

# set the order for "Ast-0", "Ast-2",  "Ast-4"
Lee_Cd_Ast_FinalObj@active.ident <- factor(Lee_Cd_Ast_FinalObj@active.ident,levels = c("Ast-0", "Ast-2", "Ast-5"))

DimPlot(Lee_Cd_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "goldenrod2", "mediumorchid3"), label = T) 

#know all the name of the sample 
table(Lee_Cd_Ast_FinalObj@meta.data$orig.ident) 

#open a new column and assign it the element in orig. ident
Lee_Cd_Ast_FinalObj$SampleNo <- Lee_Cd_Ast_FinalObj@meta.data$orig.ident

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Lee_Cd_Ast_FinalObj@meta.data$SampleNo <- ifelse(Lee_Cd_Ast_FinalObj@meta.data$SampleNo == "A47L", "CTRL_7", Lee_Cd_Ast_FinalObj@meta.data$SampleNo) # replace all the sampleNo with "A47L"  to  "CTRL_7"

#change the Control in the Genotype to CTRL
Lee_Cd_Ast_FinalObj@meta.data$Genotype <- ifelse(Lee_Cd_Ast_FinalObj@meta.data$Genotype == "Control", "CTRL", Lee_Cd_Ast_FinalObj@meta.data$Genotype)

#.svg format
svg("Lee_Cd_UMAP.svg",  width = 5,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Cd_Ast_FinalObj, reduction = "umap", pt.size = 1.0, cols =c("coral3", "goldenrod2", "mediumorchid3"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=10)) & NoLegend()
dev.off()

# cluster by sample
svg("Lee_Cd_bySample.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Cd_Ast_FinalObj, reduction = "umap", pt.size = 1.0, label = F, group.by = "SampleNo") + theme(legend.text = element_text(size = 13)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# cluster by disease
svg("Lee_Cd_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Lee_Cd_Ast_FinalObj, reduction = "umap", pt.size = 1.0, label = F, group.by = "Genotype", cols = c("green", "midnightblue")) + theme(legend.text = element_text(size = 15)) + theme(legend.key.size = unit(0.4, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=8))
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Lee_Cd_HD_Ast_FinalObj), Lee_Cd_HD_Ast_FinalObj$SampleNo), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 7), rep("HD", 3))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)
meltData <- melt(test_data, id=c("Genotype"))

# reformat the column in the melta dataframe
meltData$Genotype <- ifelse(meltData$variable == "Ast-5" & meltData$Genotype == "HD", paste0(meltData$Genotype, "5"),
                      ifelse(meltData$variable == "Ast-5" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "5"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "HD", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-2" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "2"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "HD", paste0(meltData$Genotype, "0"),
                      ifelse(meltData$variable == "Ast-0" & meltData$Genotype == "CTRL", paste0(meltData$Genotype, "0"),
                      meltData$Genotype))))))

#meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "HD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-2", "Ast-5"))


#cell distribution and sample distribution
svg("Lee_Cd_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c( "green", "midnightblue")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ Genotype, data = meltData)
summary(res.aov_PercentCell)
#            Df Sum Sq Mean Sq F value Pr(>F)    
#Genotype     5  26999    5400   211.7 <2e-16 ***
#Residuals   24    612      26                   
#---
#Signif. codes:  0 *** 0.001 ** 0.01 * 0.05 . 0.1   1

TukeyHSD(res.aov_PercentCell, conf.level = 0.99)
# all combination is significant except 
#                      diff         lwr        upr     p adj
# HD0 vs CTRL0    -5.9363318 -19.1781643   7.305501 0.5427392
# CTRL5 vs CTRL2  -7.0473710 -17.3044503   3.209708 0.1331713
# HD2 vs CTRL2     5.2534267  -7.9884057  18.495259 0.6630199
# HD5 vs CTRL2    -6.3644659 -19.6062984   6.877367 0.4687795
# HD5 vs CTRL5     0.6829051 -12.5589274  13.924738 0.9999543
# HD5 vs HD2     -11.6178927 -27.2858401   4.050055 0.0887323

# Cell distribution in each disease condition
svg("Lee_Cd_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "goldenrod2", "mediumorchid3"))  + labs(fill='Cluster')
dev.off()

#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Lee_cdHD_Astrocyte_dim15_res0.20_Vln.png",      
    width = 4*700,        # 5 x 300 pixels
    height = 4*350,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Lee_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "goldenrod2", "gray47"))& 
   theme(plot.title = element_text(size = rel(3)), axis.title.x=element_blank(), axis.title.y=element_text(size=20), axis.text.x=element_text(angle=45, hjust=1, size=31), axis.text.y=element_text( size=25)) 
dev.off()

#.svg format
svg("Lee_cdHD_Astrocyte_dim15_res0.20_Vln.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Cd_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "goldenrod2", "mediumorchid3"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()


# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("Lee_cdHD_Astrocyte_dim15_res0.20_DotPlot.png",      
    width = 9*250, height = 3*190, res = 300, pointsize = 5)   
DotPlot(Lee_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Lee_cdHD_Astrocyte_dim15_res0.20_DotPlot.svg", width = 7,  height =2,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Lee_Cd_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+ theme(axis.text.y=element_text(size=20), axis.text.x=element_blank(), legend.title = element_text(face = "bold", size = 9)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="none") + RotatedAxis()
dev.off()


#vlnplot for the number of genes, umi, mito genes detected in the nuclei and the number of nuclei showing the median value
png("Lee_cdHD_Astrocyte_dim15_res0.20_metric_genes.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lee_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lee_cdHD_Astrocyte_dim15_res0.20_metric_genes.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Ast_FinalObj, features = c("nFeature_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Lee_cdHD_Astrocyte_dim15_res0.20_metric_umi.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lee_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lee_cdHD_Astrocyte_dim15_res0.20_metric_umi.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Ast_FinalObj, features = c("nCount_RNA"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "UMIs per Nuclei")  + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.7))) + theme(legend.position="none")
dev.off()

png("Lee_cdHD_Astrocyte_dim15_res0.20_metric_mtper.png", width = 9*150, height = 10*200, res = 300, pointsize = 5)
VlnPlot(Lee_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 16), axis.text.y = element_text(size = 16), text = element_text(face = "bold")) + theme(legend.position="none")
dev.off()
# .svg format
svg("Lee_cdHD_Astrocyte_dim15_res0.20_metric_mtper.svg", width = 6,  height =6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Lee_Ast_FinalObj, features = c("percent.mt"), ncol = 1,  pt.size = 0, cols = c("coral3", "goldenrod2", "gray47")) + stat_summary(fun = median, geom='point', size = 10, colour = "black", shape = 95) + labs(title = "% Mitochondrial Genes per Nuclei") + theme(axis.title.x = element_blank()) + theme(axis.text.x = element_text(size = 18), axis.text.y = element_text(size = 16), text = element_text(face = "bold"), plot.title = element_text(size = rel(1.6))) + theme(legend.position="none")
dev.off()


```

#Gabitto SEA data
Gabitto_Ast_FinalObj <- readRDS("C:/ibrahim/Astrocyte_Psuedotime_analysis/Gabitto_MTG_Astrocyte_integration_dim15_res0.25_FinalObj_RenamedCluster.rds")
 
#plot
p1 <- DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 0.5, cols =c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"), label = F) + theme(legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.6, 'cm'))
png("Gabitto_mtgAD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()

Gabitto_SEA_AD_Astrocyte_dim15_res0.25_FinalObj$Identity <- Ident(Gabitto_SEA_AD_Astrocyte_dim15_res0.25_FinalObj)

#open another column in the metadata and rename the dementia to be AD and normal to be control
Gabitto_SEA_AD_Astrocyte_dim15_res0.25_FinalObj@meta.data <- mutate(Gabitto_SEA_AD_Astrocyte_dim15_res0.25_FinalObj@meta.data,  Conditions =ifelse(Gabitto_SEA_AD_Astrocyte_dim15_res0.25_FinalObj@meta.data$disease == "dementia", "DEM", "CTRL"))

Gabitto_Ast_FinalObj <- Gabitto_MTG_Ast_FinalObj

#renamed cluster Ast-0, Ast-2, Ast-5
new.cluster.ids <- c( "Ast-0", "Ast-1", "Ast-2",  "Ast-Int", "Ast-6", "Ast-7")
names(new.cluster.ids) <- levels(Gabitto_MTG_Ast_FinalObj)
Gabitto_MTG_Ast_FinalObj <- RenameIdents(Gabitto_MTG_Ast_FinalObj, new.cluster.ids)

# set the order for "Ast-0", "Ast-1", "Ast-Int" "Ast-2", "Ast-6", "Ast-7"
Gabitto_MTG_Ast_FinalObj@active.ident <- factor(Gabitto_MTG_Ast_FinalObj@active.ident,levels = c("Ast-0", "Ast-1", "Ast-Int", "Ast-2", "Ast-6", "Ast-7"))

DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 1.5, cols =c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"), label = T)

#Replace All Matching Rows in orig.ident in metadata with a Single Value
Gabitto_MTG_Ast_FinalObj@meta.data$orig.ident <- ifelse(Gabitto_MTG_Ast_FinalObj@meta.data$orig.ident == "merged_H20.33.025_008", "H20.33.008_025", Gabitto_MTG_Ast_FinalObj@meta.data$orig.ident) # replace all the sampleNo with "A47L"  to  "CTRL_7"

# set the order for CTRL, AD
Gabitto_MTG_Ast_FinalObj@meta.data$Condition <- factor(Gabitto_MTG_Ast_FinalObj@meta.data$Condition,levels = c("CTRL", "AD"))

#know all the name of the sample 
table(Gabitto_MTG_Ast_FinalObj@meta.data$orig.ident) 

#.svg format
svg("Gabitto_mtgAD_UMAP.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 6, family = "Arial", antialias = "gray")
DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 0.5, cols =c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"), label = F) + theme(axis.text.y=element_text(size=10), axis.text.x=element_text(size=10)) + theme(axis.title.x = element_text(size=10), axis.title.y = element_text(size=13)) & NoLegend()
dev.off()

png("Gabitto_mtgAD_UMAP.png",    
  width = 9*350,        # 5 x 300 pixels
  height = 8*400,
  res = 400,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 1, cols =c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"), label = F) + theme(axis.text.y=element_text(size=18), axis.text.x=element_text(size=18)) + theme(axis.title.x = element_text(size=18), axis.title.y = element_text(size=18)) & NoLegend()
dev.off()

# cluster by sample
svg("Gabitto_MTG_bySample.svg", width = 7,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")

png("Gabitto_MTG_bySample.png",    
  width = 9*550,        # 5 x 300 pixels
  height = 8*400,
  res = 400,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "orig.ident") + theme(legend.text = element_text(size = 5)) + theme(legend.key.size = unit(0.01, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=15), axis.text.x=element_text(size=15)) + theme(axis.title.x = element_text(size=15), axis.title.y = element_text(size=15))
dev.off()

# cluster by sample
svg("Gabitto_MTG_bySample2.svg", width =6,  height = 3,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "orig.ident") + theme(legend.text = element_text(size = 9)) + theme(legend.key.size = unit(0.01, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=8), axis.text.x=element_text(size=8)) + theme(axis.title.x = element_text(size=8), axis.title.y = element_text(size=12))
dev.off()

# cluster by disease
svg("Gabitto_MTG_byGenotype.svg", width = 5,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")

png("Gabitto_MTG_byGenotype.png",    
  width = 8*450,        # 5 x 300 pixels
  height = 8*400,
  res = 400,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
DimPlot(Gabitto_MTG_Ast_FinalObj, reduction = "umap", pt.size = 0.3, label = F, group.by = "Condition", cols = c("green", "red")) + theme(legend.text = element_text(size = 28)) + theme(legend.key.size = unit(0.7, 'cm')) + theme(plot.title = element_blank(), axis.text.y=element_text(size=15), axis.text.x=element_text(size=15)) + theme(axis.title.x = element_text(size=15), axis.title.y = element_text(size=15))
dev.off()

# Do the statistical differences in cell distribution between the genotypes using One-way ANOVA with 99% confidence level
pct_cells <- prop.table(table(Idents(Gabitto_MTG_Ast_FinalObj), Gabitto_MTG_Ast_FinalObj$orig.ident), margin = 2) *100
test_data <- as.data.frame(t(as.data.frame.matrix(pct_cells)))

#test_data$Ast2 <- test_data$"1"
#test_data$"1" <- NULL
Genotype <- c(rep("CTRL", 42), rep("AD", 41))
test_data <- cbind(test_data, Genotype)

# code worked, to display the percentage of cells in each cluster
library(reshape)

meltData <- melt(test_data, id=c("Genotype"))
meltData$Genotype <- factor(meltData$Genotype,levels = c("CTRL", "AD"))
meltData$variable <- factor(meltData$variable,levels = c("Ast-0", "Ast-1", "Ast-Int", "Ast-2", "Ast-6", "Ast-7"))

#cell distribution and sample distribution
svg("Gabitto_MTG_Ast_Sample_statistics_CellProportion_ScaterPlot.svg", width = 7,  height = 5,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(meltData, aes(x= Genotype, y=value, group = Genotype, color = Genotype, fill = Genotype)) +
    scale_color_manual(values=c( "green", "red")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=5, alpha=0.8) + 
  labs(x = "Genotype", y = "% distribution of astrocyte") + facet_wrap( variable ~ ., ncol=4, scales="free_y") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#significant differences in the distribution of cells between the samples
res.aov_PercentCell <- aov(value ~ Genotype, data = meltData)
summary(res.aov_PercentCell)
TukeyHSD(res.aov_PercentCell, conf.level = 0.99)

# Cell distribution in each disease condition
svg("Gabitto_MTG_Ast_Cell_Distribution_In_Each_Disease_Group.svg", width = 7,  height = 6,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
ggplot(meltData, aes(fill=variable, y=value, x= Genotype)) + 
    geom_bar(position="fill", stat="identity") +   theme(axis.text=element_text(size=20), axis.title.x =element_blank(), axis.title.y = element_text(size=24)) + ylab("Proportion of astrocyte subpopulation") + theme(legend.title= element_blank(), legend.text = element_text(size = 20)) + theme(legend.key.size = unit(0.4, 'cm')) +
    scale_fill_manual(values = c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"))  + labs(fill='Cluster')
dev.off()

#unique markers
markerAst_vln <- c('GPC5', 'GJA1', 'VCAN') 

png("Gabitto_mtgAD_Astrocyte_dim15_res0.25_Vln.png",      
    width = 4*700,        # 5 x 300 pixels
    height = 4*350,
    res = 300,            # 300 pixels per inch
    pointsize = 3)        # smaller font size
VlnPlot(Gabitto_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"))& 
   theme(plot.title = element_text(size = rel(2.5)), axis.title.x=element_blank(), axis.title.y=element_text(size=18), axis.text.x=element_text(angle=45, hjust=1, size=17), axis.text.y=element_text( size=20)) 
dev.off()

#.svg format
svg("Gabitto_mtgAD_Astrocyte_dim15_res0.25_Vln.svg", width = 8,  height = 4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
VlnPlot(Gabitto_MTG_Ast_FinalObj, assay = "RNA", features = markerAst_vln, pt.size = 0, ncol =3, cols = c("coral3", "deepskyblue3", "dodgerblue4", "goldenrod2", "gold4", "forestgreen"))& 
  theme(plot.title = element_text(size = rel(2)), axis.title.x=element_blank(), axis.title.y=element_text(size=13), axis.text.x=element_text(angle=45, hjust=1, size=19), axis.text.y=element_text( size=13)) 
dev.off()

# Dot plot
# markers for all cell types in reverse order
marker_genes <- c( "GPC5", "NRXN1", "TRPM3", "APOE", "CST3", "FTL", "CKB", "CLU", "ITM2C", "CPE", "TUBB2B", "VIM", "GLUL", "PSAP", "GJA1", "AGT", "ENO1", "DPP10", "GFAP", "PLEKHA5", "VCAN", "KAZN")  

png("_Astrocyte_dim15_res0.25_DotPlot.png",      
    width = 9*250, height = 3*250, res = 300, pointsize = 5)   
DotPlot(Gabitto_Ast_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8)+  theme(legend.title = element_text(face = "bold", size = 8)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.text = element_text(size = 6)) + theme(legend.key.size = unit(0.2, 'cm')) + RotatedAxis()
dev.off()

#.svg format
svg("Gabitto_mtgAD_Astrocyte_dim15_res0.25_DotPlot2222.svg", width = 9,  height =4,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")
DotPlot(Gabitto_SEA_AD_Astrocyte_dim15_res0.25_FinalObj, assay = "SCT", features = marker_genes, dot.scale = 8) & theme(axis.text.y=element_text(size=20), axis.text.x=element_text(size=15, angle = 55, hjust=1), legend.title = element_text(face = "bold", size = 15)) + theme(axis.title.x = element_blank(), axis.title.y = element_blank()) + theme(legend.position="bottom") + RotatedAxis()
dev.off()

```{r}
#Draw barplot for metric of each dataset
```

#get quality metrix
load("~/ibrahim/snRNA_seq_DEG_analysis/GO_DEG_analysis/2024_06_24_Ast_Universal_genes_for_GO_analysis/2024_06_24_All_dataset_FinalObj_and_universal_gene.RData")

obj_list <- list(GA_ec = Grubman_Ast_FinalObj, FR_acc = Feleke_LBD_Ast_FinalObj, LS_pfc = Lau_Ast_FinalObj, XJ_pu = Xu_Ast_FinalObj, ZY_oc = Zhou_NHD_Ast_FinalObj, LH_pu= Lee_Pu_HD_Ast_FinalObj, Lee_cd = Lee_Cd_HD_Ast_FinalObj, GM_mtg = Gabitto_MTG_Ast_FinalObj)

#nFeature
median_genes_df <- do.call(rbind, lapply(names(obj_list), function(dataset_name) {
  obj <- obj_list[[dataset_name]]
  
  # Replace 'orig.ident' with your sample ID column name if different
  df <- data.frame(
    sample = obj$orig.ident,
    nFeature_RNA = obj$nFeature_RNA
  )
  
  # Calculate median nFeature_RNA per sample
  median_per_sample <- df %>%
    group_by(sample) %>%
    summarise(median_nFeature_RNA = median(nFeature_RNA)) %>%
    mutate(dataset = dataset_name)
  
  median_per_sample
})) 

# Reorder columns if you like
median_genes_df <- median_genes_df[, c("dataset", "sample", "median_nFeature_RNA")]
median_genes_df$dataset <- factor(median_genes_df$dataset, levels = unique(median_genes_df$dataset))


#nCount
median_umis_df <- do.call(rbind, lapply(names(obj_list), function(dataset_name) {
  obj <- obj_list[[dataset_name]]
  
  # Replace 'orig.ident' with your sample ID column name if different
  df <- data.frame(
    sample = obj$orig.ident,
    nCount_RNA = obj$nCount_RNA
  )
  
  # Calculate median nCount_RNA per sample
  median_per_sample <- df %>%
    group_by(sample) %>%
    summarise(median_nCount_RNA = median(nCount_RNA)) %>%
    mutate(dataset = dataset_name)
  
  median_per_sample
})) 

# Reorder columns if you like
median_umis_df <- median_umis_df[, c("dataset", "sample", "median_nCount_RNA")]
median_umis_df$dataset <- factor(median_umis_df$dataset, levels = unique(median_umis_df$dataset))


#total nuclei
total_nuclei_df <- do.call(rbind, lapply(names(obj_list), function(dataset_name) {
  obj <- obj_list[[dataset_name]]
  
  df <- data.frame(
    sample = obj$orig.ident  # change if sample ID column differs
  )
  
  total_per_sample <- df %>%
    group_by(sample) %>%
    summarise(total_nuclei = n()) %>%
    mutate(dataset = dataset_name)
  
  total_per_sample
}))

# Reorder columns
total_nuclei_df <- total_nuclei_df[, c("dataset", "sample", "total_nuclei")]
total_nuclei_df$dataset <- factor (total_nuclei_df$dataset, levels = unique(total_nuclei_df$dataset))


#Gene per nuclei
svg("Dataset_Gene_Metric_ScaterPlot.svg", width = 6,  height = 8,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(median_genes_df, aes(x= dataset, y=median_nFeature_RNA, group = dataset, color = dataset, fill = dataset)) +
    scale_color_manual(values=c("blue4", "green", "blueviolet", "lightslateblue", "goldenrod", "pink", "cadetblue",  "red2")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "Gene per Nuclei") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=29), axis.text.y=element_text(size=23)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#Umi per nuclei
svg("Dataset_UMI_Metric_ScaterPlot.svg", width = 6,  height = 8,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(median_umis_df, aes(x= dataset, y=median_nCount_RNA, group = dataset, color = dataset, fill = dataset)) +
    scale_color_manual(values=c("blue4", "green", "blueviolet", "lightslateblue", "goldenrod", "pink", "cadetblue",  "red2")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "UMI per Nuclei") +
  theme(legend.position="none", axis.text=element_text(size=18.5), axis.title.x =element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=29), axis.text.y=element_text(size=23)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()

#total nuclei per sample
#Umi per nuclei
svg("Dataset_total_nuclei_Metric_ScaterPlot.svg", width = 6,  height = 8,  onefile = TRUE, pointsize = 8, family = "Arial", antialias = "gray")  
ggplot(total_nuclei_df, aes(x= dataset, y=total_nuclei, group = dataset, color = dataset, fill = dataset)) +
    scale_color_manual(values=c("blue4", "green", "blueviolet", "lightslateblue", "goldenrod", "pink", "cadetblue",  "red2")) +
  geom_point(position=position_jitter(w=0.1,h=0), size=4, alpha=0.8) + 
  labs(x = "Genotype", y = "Number of Nuclei") +
  theme(legend.position="none", axis.text=element_text(size=26), axis.title.x =element_blank(), axis.text.x = element_text(angle = 45, vjust = 1, hjust=1), axis.title.y = element_text(size=29), axis.text.y=element_text(size=23)) +
  theme(strip.text.x = element_text(size=19, face = "bold"))
  dev.off()
        
# ###############################################################
saveRDS(Lau_PFC_Ast_FinalObj, file = "Lau_Astrocyte_ClusterRenamed_dim15_res0.25_ObjFinal.rds")
saveRDS(Grubman_EC_Ast_FinalObj, file = "Grubman_Astrocyte_ClusterRenamed_dim15_res0.25_ObjFinal.rds")

saveRDS(Feleke_ACC_Ast_FinalObj, file = "Feleke_LBD_Astrocyte_MergedClusterRenamed_dim15_res0.25_CleanObj_Final.rds")
saveRDS(Xu_Pu_Ast_FinalObj, file = "Putamen_Astrocyte_FinalObj_dim15_res0.25_ClusterRenamed.rds")

saveRDS(Lee_Pu_Ast_FinalObj, file = "Lee_Pu_HD_Astrocyte_dim15_res0.25_FinalObj.rds")
saveRDS(Lee_Cd_Ast_FinalObj, file = "Lee_Cd_HD_Astrocyte_dim15_res0.25_FinalObj.rds")

saveRDS(Zhou_OC_Ast_FinalObj, file = "Zhou_OC_NHD_Astrocyte_integrated_dim15_res0.25_RenamedCluster.rds")
saveRDS(Gabitto_MTG_Ast_FinalObj, file = "Gabitto_MTG_Astrocyte_integration_dim15_res0.25_FinalObj_RenamedCluster.rds")

