---
title: "Human_2023_Zhou_OC_NHD_Colona_Astrocyte_Subset_Analysis"
author: "Ibrahim Saliu"
output: html_document
---
# set working directory
setwd("C:/ibrahim/Human_2023_Zhou_OC_NHD_Colona/Human_2023_Zhou_OccipitalCortex_NHD_Astrocyte")

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library
library('Seurat')
library(sctransform)
library(cowplot)
library(gplots) # for heatmap
library(RColorBrewer) # colorRampPalette(), brewer.pal
library(ggplot2) # qplot(), position_nudge(), geom_text()
library(dplyr)
library("limma")
library(VennDiagram)
library(reshape)
library("ggpubr")
library(UpSetR)

```

#read in the astrocyte data
Astrocyte_Original_Obj <- readRDS("C:/ibrahim/Human_2023_Zhou_OC_NHD_Colona/Human_2023_Zhou_OccipitalCortex_NHD_FullData_FinalObj_Astrocyte.rds")

# plot the UMAP
DimPlot(Astrocyte_Original_Obj, reduction = "umap", label = TRUE)

#check the number of cells in each cluster
table(Astrocyte_Original_Obj@active.ident)
#Astrocyte 
#     5807

# split the object cluster and integrate
DefaultAssay(Astrocyte_Original_Obj) <- "RNA"
Astrocyte_Original_Obj[["integrated"]] <- NULL
Astrocyte_Original_Obj[["SCT"]] <- NULL

obj.list <- SplitObject(Astrocyte_Original_Obj, split.by = "orig.ident")

sapply(obj.list, function(x) length(x@meta.data$Genotype) )

#make sample statistics
nCells <- sapply(obj.list, function(x) table(x@meta.data$orig.ident) )
medianFeature <- sapply(obj.list, function(x) median(x@meta.data$nFeature_RNA) )
medianCount <- sapply(obj.list, function(x) median(x@meta.data$nCount_RNA) )

SampleStatistics <- rbind(nCells, medianCount)
SampleStatistics <- rbind(SampleStatistics, medianFeature)

write.csv(SampleStatistics, file = "SampleStatistics_before_sample_merge.csv")

#normalization using SCTransform approach
for (i in 1:length(obj.list)) {
    obj.list[[i]] <- SCTransform(obj.list[[i]],  vars.to.regress = c("nCount_RNA"),verbose = FALSE)
}

grep(pattern = "^MT-", x= rownames(Astrocyte_Original_Obj@assays$RNA), value = TRUE)

```{r}

# Identification of top consistent highly variable genes across the samples
alldata.features <- SelectIntegrationFeatures(object.list = obj.list, nfeatures = 3000)

alldata_prepsct <- PrepSCTIntegration(object.list = obj.list, anchor.features = alldata.features, verbose = FALSE)

# determine the inter-sample anchors for integration
alldata.anchors <- FindIntegrationAnchors(object.list = alldata_prepsct, normalization.method = "SCT", reduction = "cca", dim = 1:30, anchor.features = alldata.features,  verbose = FALSE)

#Integrate the data
data_integrated <- IntegrateData(anchorset = alldata.anchors, normalization.method = "SCT", verbose = FALSE, dims = 1:30)

# Dimensionality Reduction with PCA & UMAP, then clustering
data_integrated <- RunPCA(data_integrated,  npcs = 30, verbose = FALSE)
dims=1:30
data_integrated <- RunUMAP(data_integrated, reduction = "pca", dims = dims)
data_integrated <- FindNeighbors(data_integrated, reduction = "pca", dims = dims)

data_integrated <- FindClusters(data_integrated, resolution = 1.2)

# Visualise the UMAP
p1 <- DimPlot(data_integrated, reduction = "umap", group.by = "orig.ident")
p2 <- DimPlot(data_integrated, reduction = "umap", label = TRUE)
png("Zhou_OC_NHD_Astrocyte_integration_dim30_res1.2.png",    
  width = 10*500,        # 5 x 300 pixels
  height = 5*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1 + p2
dev.off()

# find conserved marker gene for each cluster
DefaultAssay(data_integrated) <- "SCT"

Conserved_cluster_markers <- list() # list starting subscript is 1, can't use 0
for (i in 0:(nlevels(data_integrated@active.ident)-1)) {
  Conserved_cluster_markers[[i+1]] <- FindConservedMarkers(data_integrated, assay = "SCT", slot = "data", test.use = "wilcox", min.pct = 0.2, ident.1 = i, grouping.var = "Genotype", verbose = FALSE, only.pos = TRUE, recorrect_umi = FALSE)
  
   # apply cutoff
    Conserved_cluster_markers[[i+1]] <- Conserved_cluster_markers[[i+1]][ which(Conserved_cluster_markers[[i+1]]$max_pval < 0.05), ]
}

#output marker genes for each cluster
for (n in 1:length(Conserved_cluster_markers)) {
  file_number = n-1
  outputfile <- paste("Zhou_OC_NHD_Astrocyte_dim30_res1.2_ConservedMarkerGenes_cluster_", file_number,".csv", sep="")
  write.csv(Conserved_cluster_markers[n], outputfile)
}

# plot conserved marker genes as feature and violin plot, first normalize the data
DefaultAssay(data_integrated) <- "RNA"
data_integrated <- NormalizeData(data_integrated, verbose = FALSE)

# plot conserved marker genes as feature plot
for (n in 1:length(Conserved_cluster_markers)) {
  file_number = n-1
  outputfile <- paste("Zhou_OC_NHD_Astrocyte_dim30_res1.2_ConservedMarkerGenes_cluster_", file_number,"_top20.png", sep="")
  
  marker_genes <- row.names(Conserved_cluster_markers[[n]][1:20,])
  png(outputfile, width = 16*300, height = 13*300, res = 300, pointsize = 5)     
  print(FeaturePlot(data_integrated, features = marker_genes, min.cutoff = "q9", pt.size = 0.3, max.cutoff = "q90", ncol = 4))
  dev.off()
}

# plot conserved marker genes as vln plot
for (n in 1:length(Conserved_cluster_markers)) {
  file_number = n-1
  outputfile <- paste("Zhou_OC_NHD_Astrocyte_dim30_res1.2_ConservedMarkerGenes_vln_cluster_", file_number,"_top20.png", sep="")
  
  marker_genes <- row.names(Conserved_cluster_markers[[n]][1:20,])
  png(outputfile, width = 16*300, height = 13*300, res = 300, pointsize = 5)     
  print(VlnPlot(data_integrated, assay = "RNA", features = marker_genes) )
  dev.off()
}

#do the violin plot to see the quality of all the clusters
png("Zhou_OC_NHD_Astrocyte_dim30_res1.2_quality.png", width = 16*300, height = 13*300, res = 300, pointsize = 5)
VlnPlot(data_integrated, features = c("nFeature_RNA", "nCount_RNA", "percent.mt"), ncol = 3)
dev.off()

table(data_integrated@active.ident)

```

# Rename cluster
new.cluster.ids <- c( "Ast-0", "Ast-1", "Ast-2") 
names(new.cluster.ids) <- levels(data_integrated)
data_integrated <- RenameIdents(data_integrated, new.cluster.ids)

# set the order for Genotype 
data_integrated@meta.data$Genotype <- factor(data_integrated@meta.data$Genotype,levels = c("CTRL", "NHD"))

# set the order for active.ident "Ast-0", "Ast-1",  "Ast-2", "Ast-2.2"
data_integrated@active.ident <- factor(data_integrated@active.ident,levels = c("Ast-0", "Ast-1", "Ast-2"))

p1 <- DimPlot(data_integrated, reduction = "umap", pt.size = 1.5, cols =c("coral3", "deepskyblue3", "goldenrod2"), label = F) + theme(legend.text = element_text(size = 20))
png("Zhou_OC_NHD_Astrocyte_dim15_res0.25_RenamedCluster.png",    
  width = 5*400,        # 5 x 300 pixels
  height = 4*400,
  res = 300,            # 300 pixels per inch
  pointsize = 8)        # smaller font size
  p1
dev.off()


saveRDS(data_integrated, file = "Zhou_OC_NHD_Astrocyte_integrated_dim15_res0.25_RenamedCluster.rds")
save.image("Zhou_OC_NHD_Astrocyte_integrated_dim15_res0.25_RenamedCluster.RData")


```{r}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
